### 2023-05-28 documentation/article.tex
Second commit

\input{preamble.tex}

\begin{document}
	\title{A tutorial of Solving the defined integral problem with OpenMp}
	\author{André Furlan - UNESP - Universidade Estadual Paulista "Júlio de Mesquita Filho"}
	\date{2023-06-01}
	\maketitle

	\begin{abstract}
		The Trapezoid rule algorithm used in this tutorial is \textbf{not} the most fast method to calculate an interval defined integral but is a good example of how to make programs runs quicker by using OpenMp for parallelism. Hopefully it will help the understanding of how this library works and how to use it.
	\end{abstract}

	\begin{IEEEkeywords}
		OpenMp, Paralelism, C++, C, Tutorial
	\end{IEEEkeywords}

	\section{Introduction}
	\par In High Performance Computation (HPC) the goal (other than solve a problem) is to use most of the computation power of the machines involved. Most of the time this means use all processors an its cores at maximum, avoiding idle intervals.
	\par By using the OpenMp \cite{openmp08} libraries it is possible, using \textbf{\#pragma} compiler directives, turn a sequential program into a parallel one. The problem in focus here is the defined interval integral calculation. To solve it the program must calculate the area bellow the plot by summing up the area of several trapezoids \cite{hildebrand1987introduction}. The more trapezoids fitted, the more precise are the values obtained. And it is here that the parallelism enters: By dividing the work of area calculation in lots of threads its possible to diminish the effective time of execution.
	\section{The trapezoid rule}
	\par Be $a$ a point that marks the start of a interval in which the calculation must be done and $b$ its end, than an \textbf{approximation} of the integral value is given by equation \ref{eq:firstApproximation}. In figure \ref{fig:trapezoidalruleillustration} the basic idea is illustrated.\newline

	\begin{equation}
		\int^a_b{f(x) dx} \approx (b-a).\dfrac{1}{2}(f(a)+f(b))
		\label{eq:firstApproximation}
	\end{equation}

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.7\linewidth]{images/Trapezoidal_rule_illustration}
		\caption[Approximation of an integral]{In pink the approximation(trapezoid area), in blue the actual function curve. $f(a)$ and $f(b)$ are the trapezoid bases and the interval $a, b$ its height. }
		\label{fig:trapezoidalruleillustration}
	\end{figure}

	\par The figure \ref{fig:trapezoidalruleillustration} and the equation \ref{eq:firstApproximation} shows a pretty coarse approximation! What if the $[a, b]$ interval get subdivided (partitioned) like in figure \ref{fig:trapezoidalruleillustration2}? Well, then it is possible to get a value closer to the real one!\newline

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.7\linewidth]{images/Trapezoidal_rule_illustration2}
		\caption[Enhanced approximation of an integral]{In pink the \textbf{enhanced} approximation(trapezoid area), in blue the actual function curve. $f(a), f(b), f(c), f(d), f(e)$ are the trapezoid bases and the intervals $[a, d],[d,c], [c,e], [e,b]$ its heights. Note that now there are 4 trapezoids i.e. 4 areas that must be calculated. }
		\label{fig:trapezoidalruleillustration2}
	\end{figure}

	\par To obtain the result all areas must be summed up, therefore the equation \ref{eq:firstApproximation} turns into \ref{eq:secondApproximation}. So, as many partitions are made, the more sums are added.\newline

	\begin{equation}
		\begin{aligned}
			\int^a_b{f(x) dx} \approx 
			&(d-a).\dfrac{1}{2}(f(d)+f(a))+\\
			&(c-d).\dfrac{1}{2}(f(c)+f(d))+\\
			&(e-c).\dfrac{1}{2}(f(e)+f(c))+\\
			&(b-e).\dfrac{1}{2}(f(b)+f(e))
		\end{aligned}
		\label{eq:secondApproximation}
	\end{equation}

	\par Let $x = [a,b]$ and $n=partitions+1=points$ of $[a,b]$ than it is true to say that $x_0 = a < x_1 < ... < x_{n-1} < x_n = b$. Being the interval between two points $\Delta_{x_k} = x_k - x_{k-1}$ than the equation \ref*{eq:secondApproximation} can be rewrote as in the equation \ref{eq:genericApproximation}.\newline

	\begin{equation}
		\int^a_b{f(x) dx} \approx \sum_{k=1}^{n} \dfrac{f(x_{k-1})+f(x_k)}{2} . \Delta_{x_k}
		\label{eq:genericApproximation}
	\end{equation}

	\par And this is (finally) the idea that will be implemented ahead!

	\section{Implementation}
		\par First of all it is necessary to define a function to integrate, the one used here is defined in equation \ref{eq:function} and the integration interval begins in \textbf{0 until 100}. The performance will be tested using \textbf{1, 2, 4 and 8 threads}.  Note that this method supports any function, this one was chosen for example purposes only.\newline

		\begin{equation}
			f(x) = \sqrt{100^2 - x^2}
			\label{eq:function}
		\end{equation}

		\par Then, for the impatient ones, here is the complete code that will be discussed:
		\lstinputlisting[language=C++]{code/integral.cpp}

	\section{Discussion}
		\par This section focus only in the OpenMp directives, its assumed that the reader knows a minimum of C++.\newline

		\par The lines 5 to 7 includes OpenMp header files if the build system used detected such library.
		\begin{lstlisting}[language=C++]
#ifdef _OPENMP
	#include <omp.h>
#endif
		\end{lstlisting}

		\par From line 9 to 12, the function to be integrated is defined. Note the \textit{inline} statement: Inlinning functions \textbf{may} avoid the intrinsic overhead of calling a function but, since this it is not a command, the compiler may ignore this request.
		\begin{lstlisting}[language=C++]
inline long double f(long double x)
{
	return std::sqrt(std::pow(100,2) - std::pow(x,2));
}
		\end{lstlisting}

		\par Line 22 states that the variable \textit{sum} is shared by all threads all the other variables are not.
		\begin{lstlisting}[language=C++]
#pragma default(none) shared(soma)
		\end{lstlisting}

		\par At the 24 position OpenMp starts a parallel zone: From now on the instructions are replicated in several  threads according to parallelization rules defined in the \textbf{operational system ambient variables} \cite{openmp08SV}. The zones are defined using curling braces (same as blocks in C++).
		\begin{lstlisting}[language=C++]
#pragma omp parallel
		\end{lstlisting}

		\par The most important part is declared here in line 26. This directive is the heart of the parallelization for this algorithm. For each thread \textit{sum} is initialized to 0 then, after all calculations, the obtained values are summed together. For this directive it have to be possible to known beforehand the number of iterations in execution time i.e. the number of iterations cannot change while the loop are being executed. Also the k variable (used in for loop) must \textbf{not} be modified within the loop otherwise it can be changed like in k++ statement. Since, in this specific case, a summation is need OpenMp have a convenient \textit{reduction} directive that enables it. The \textit{reduction} directive can be used with other types of operations like \textit{subtractions}, \textit{maximums} and \textit{minimums}.

		\begin{lstlisting}[language=C++]
#pragma omp for reduction(+ : sum)
		\end{lstlisting}

		\par Finally, at the line 31, a barrier to the execution of the program is created in order to ensure that all threads has been finished until this point. Now it is possible to do another operations with the outputted sum.  This is optional in this \textbf{specific case} because there is an implicit barrier in \textit{for} directive. It is important to note that \textbf{there is no code} associated with this directive i.e. it just marks a point where all threads must arrive. \textbf{Warning!} All or none of the threads must encounter the barrier otherwise a \textbf{deadlock} happens.

		\begin{lstlisting}[language=C++]
#pragma omp barrier
		\end{lstlisting}

		\par Among the aforementioned \textbf{operational system ambient variables} one is the most important for the sake of this program: OMP\_NUM\_THREADS. This variable sets the number of threads (an integer) that are going to be used when running the algorithm. If no value is assigned than the maximum of threads supported are used.

	\section{Performance tests}
		\par The tests were done using an \textit{AMD Ryzen 5 5000 series} running an \textit{Gnu/Linux operational system} with \textit{30GB of RAM}.\newline

		\par The tables bellow are shows the results according to the number of threads.

		\begin{table}[h]
			\caption{export OMP\_NUM\_THREADS=1}
			\begin{center}
				\begin{tabular}{|l|l|l|l|}
					\hline
					Interval $\Delta_{x_k}$  &  Partitions  &  Time taken in sec &  Result \\
					\hline
					0.000001  &  100000000  &  0.670965  &  7853.981634 \\
					\hline
					0.000010  &  10000000  &  0.067101  &  7853.981633 \\
					\hline
					0.000100  &  1000000  &  0.006715  &  7853.981617 \\
					\hline
				\end{tabular}
			\end{center}
		\end{table}

		\begin{table}[h]
			\caption{export OMP\_NUM\_THREADS=2}
			\begin{center}
				\begin{tabular}{|l|l|l|l|}
					\hline
					Interval $\Delta_{x_k}$  &  Partitions  &  Time taken in sec &  Result \\
					\hline
					0.000001  &  100000000  &  0.341421  &  7853.981634 \\
					\hline
					0.000010  &  10000000  &  0.033588  &  7853.981633 \\
					\hline
					0.000100  &  1000000  &  0.003378  &  7853.981617 \\
					\hline
				\end{tabular}
			\end{center}
		\end{table}

		\begin{table}[h]
			\caption{export OMP\_NUM\_THREADS=4}
			\begin{center}
				\begin{tabular}{|l|l|l|l|}
					\hline
					Interval $\Delta_{x_k}$  &  Partitions  &  Time taken in sec &  Result \\
					\hline
					0.000001  &  100000000  &  0.172951  &  7853.981634 \\
					\hline
					0.000010  &  10000000  &  0.016810  &  7853.981633 \\
					\hline
					0.000100  &  1000000  &  0.001684  &  7853.981617 \\
					\hline
				\end{tabular}
			\end{center}
		\end{table}

		\begin{table}[h]
			\caption{export OMP\_NUM\_THREADS=8}
			\begin{center}
				\begin{tabular}{|l|l|l|l|}
					\hline
					Interval $\Delta_{x_k}$  &  Partitions  &  Time taken in sec &  Result \\
					\hline
					0.000001  &  100000000  &  0.144975  &  7853.981634 \\
					\hline
					0.000010  &  10000000  &  0.014428  &  7853.981633 \\
					\hline
					0.000100  &  1000000  &  0.001436  &  7853.981617 \\
					\hline
				\end{tabular}
			\end{center}
		\end{table}

	\section{Conclusion}
	\par Conclusion

	\bibliographystyle{plain}
	\bibliography{references.bib}
\end{document}

### 2023-05-28 code/main.cpp

@@ -1,0 +1,110 @@
+#include <cmath>
+#include <iomanip>
+#include <iostream>
+#include <ostream>
+
+#ifdef _OPENMP
+#include <omp.h>
+#endif
+
+/*
+ * Inlinning functions so the call overhead
+ * is avoided
+ */
+inline long double f(long double x)
+{
+    return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
+}
+
+/*
+ * Inlinning the functions so the call overhead
+ * is avoided. Note that **inline** it is not a
+ * command, the compiler may **ignore** the inline
+ * request.
+ */
+inline long double integrate(long double (*f)(long double),
+                             double lowerLimit,
+                             double upperLimit,
+                             int partitions)
+{
+    double h = (upperLimit - lowerLimit) / partitions;
+    long double sum = 0;
+
+/*
+ * sum is shared by all threads
+ * all the other variables are not
+ */
+#pragma default(none) shared(soma)
+    {
+/*
+ * Starts a parallel zone, from now on the
+ * instructions are replicated in several 
+ * threads according to parallelization rules
+ * bellow.
+ */
+#pragma omp parallel
+        {
+/*
+ * For each thread sum is initialized to zero
+ * then, after all calculations, the obtained values
+ * are summed together.
+ * For this directive it have to be possible to known
+ * beforehand the number of iterations in execution time
+ * i.e. the number of iterations cannot change while the
+ * loop are being executed. Also the k variable must NOT
+ * be modified within the loop otherwise it can be changed
+ * like in k++ statement.
+ */
+#pragma omp for reduction(+ : sum)
+            for (int k = 1; k < partitions - 1; k++) {
+                sum += f(lowerLimit + k * h);
+            }
+        }
+/*
+ * Creates a barrier to the execution of the program
+ * in order to ensure that all threads has been finished
+ * until this point. Now it is possible to do 
+ * another operations with the outputted sum.
+ * This is optional in this **specific case** because
+ * there is an implicit barrier in **for** directive.
+ * It is important to note that THERE IS NO CODE associated
+ * with this directive i.e. it just marks a point where all
+ * threads must arrive.
+ * Warning! All or none of the threads must encounter the 
+ * barrier otherwise an DEADLOCK happens.
+ */
+#pragma omp barrier
+
+        sum *= 2;
[…]

### 2023-05-28 documentation/article.tex
Url to repo added
Minor fixes

@@
 Hopefully it will help the understanding of how this library works and how to use it.
+All sources can be accessed at \href{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}
 	\end{abstract}
@@
 
-		\par The tables bellow are shows the results according to the number of threads.
+		\par The tables bellow shows the results according to the number of threads.
 

### 2023-05-28 documentation/article.tex
Minor fixes

@@
 	\section{Performance tests}
-		\par The tests were done using an \textit{AMD Ryzen 5 5000 series} running an \textit{Gnu/Linux operational system} with \textit{30GB of RAM}.\newline
+		\par The tests were done using an \textit{AMD Ryzen 5 5000 series} processor running an \textit{Gnu/Linux operational system} with \textit{30GB of RAM}.\newline
 

### 2023-06-09 documentation/article01/article.tex
Beginning a new article

--- 2023-05-28 documentation/article.tex
@@
 < x_{n-1} < x_n = b$.
-Being the interval between two points $\Delta_{x_k} = x_k - x_{k-1}$ than the equation \ref*{eq:secondApproximation} can be rewrote as in the equation \ref{eq:genericApproximation}.\newline
+Being the interval between two points $\Delta_{x_k} = x_k - x_{k-1}$ than the equation \ref*{eq:secondApproximation} can be rewrote as the equation \ref{eq:genericApproximation}.\newline
 
@@
 	\section{Conclusion}
-	\par Conclusion
+	\par This is just a scratch on the surface of OpenMp please pay a visit to their website to learn more: \href{https://www.openmp.org/}{https://www.openmp.org/}.
+	\par \textit{That is all folks!}
 

### 2023-06-12 documentation/article02/article.tex
Pretty good advance on the second article.
More one night asleep. Muauauahahahahahahha.

\input{preamble.tex}

\begin{document}
	\title{A brief review of Quantum Mechanics for Computer Scientists}
	\author{André Furlan - UNESP - Universidade Estadual Paulista "Júlio de Mesquita Filho"}
	\date{2023-06-01}
	\maketitle

	\begin{abstract}
		Quantum computers are one of the most promising technologies currently. However, a lack of understanding of its basic concepts may hinder even experienced computer scientists from exploring its possibilities. This document aims to provide a brief review of these concepts to enable those interested in quantum computation to understand and utilize its ideas effectively.
	\end{abstract}

	\begin{IEEEkeywords}
		quantum mechanics, quantum computation, qubit
	\end{IEEEkeywords}

	\section{Introduction}
		\par In the same century that classical computers, with their classical bits that can only assume one of two values (0 or 1), emerged, another field of knowledge arose: Quantum Mechanics. This field was developed to explain phenomena in the microscopic domain that could not be adequately described by Newtonian mechanics. As some quantum phenomena, such as entanglement and superposition, exhibit counterintuitive behavior, scientists began exploring the possibility of utilizing these behaviors in computing, leading to the birth of quantum computing.

	\section{Quantum states}
		\par In quantum mechanics, a \textbf{state} refers to a mathematical representation that provides information about the properties of a physical system. It encapsulates the essential characteristics and behavior of the system, such as its observable quantities and their possible values.\newline

		\par The state of a quantum system is described using a formalism called \textbf{state vectors}, which belong to a \textbf{vector space} (more on this in section \ref{sec:vectorspace}) associated with the system. These vectors are often denoted using the "ket" notation (section \ref{sec:ketNotation}), introduced by physicist Paul Dirac, where a state vector is represented as $|\psi\rangle$, pronounced as "ket psi".\newline

		\par The state vector represents the system's quantum state, encoding both the information about the system's observable quantities and the probabilities of obtaining specific outcomes upon measurement. The state vector can evolve over time according to the laws of quantum mechanics, capturing the dynamics and changes in the system as it interacts with its environment.\newline

		\par A quantum state can be in a \textbf{superposition}, meaning it is not limited to a single definite value but rather exists as a combination of multiple possible states. This superposition arises due to the fundamental principle of quantum mechanics known as \textbf{superposition principle}, which allows states to be in a coherent linear combination of basis states.\newline

		\par It is important to note that a state is not a physical object itself, but rather a mathematical representation that provides a comprehensive description of the system's properties.

	\section{Complex vector space}
		\label{sec:vectorspace}
		\par A \textbf{vector space} ($V$) over the complex numbers $\mathbb{C}$ is a mathematical structure that allows for the representation of its elements (vectors) and adheres to the following set of rules:

		\begin{itemize}
			\item $\alpha \in \mathbb{C}, \alpha . |V_1\rangle \in V$
			\item $\vec{0} , \vec{0} + |V_1\rangle = |V_1\rangle$ (null vector)
			\item $|V_1\rangle + |V_2\rangle = |V_2\rangle + |V_1\rangle$ (commutative)
			\item $|V_1\rangle \in V, |V_2\rangle \in V \implies |V_1\rangle + |V_2\rangle \in V$
			\item $|-V_1\rangle; |V_1\rangle + |-V_1\rangle=\vec{0}, |-V_1\rangle=-|V_1\rangle$ 
			\item $|V_1\rangle + (|V_2\rangle + |V_3\rangle) = (|V_1\rangle + |V_2\rangle) + |V_3\rangle$ (associative)
			\item $\alpha,\beta \in \mathbb{C}; \alpha (|V_1\rangle + \beta . |V_2\rangle) = \alpha . |V_1\rangle + \alpha . \beta . |V_2\rangle $ (distributive)
		\end{itemize}

		\par It is important to highlight that when using the "BraKet" notation described in section \ref{sec:DiracNotation}, the \textbf{basis}, which are \textbf{normalized and mutually orthogonal} vectors within the vector space, is used in combination with other parameters to represent a specific state of interest.\newline

		\par In quantum mechanics these \textbf{basis vectors} are called \textbf{system possible states}.

	\section{Dirac notation}
		\label{sec:DiracNotation}

		\par In quantum mechanics, it is crucial to define the \textbf{range} of values that a particle's state can assume. This concept that will be discussed deeply in section \ref{sec:probAmpli} is known as \textbf{probability amplitude} and it employs two notations: the "ket" and the "bra". These notations are commonly referred to as the \textbf{Dirac's "BraKet" notation}.

		\subsection{Ket notation}
			\label{sec:ketNotation}
			\par The combination of symbols $|$ and $\rangle$ with a value inside is referred to as a "ket." An example is demonstrated in Equation \ref{eq:ketNotationVector}.

			\begin{equation}
				|K\rangle = \vec{K} = \{a, b, c, d\}
				\label{eq:ketNotationVector}
			\end{equation}

			\par Where $a, b, c, d \in \mathbb{Q}$ and are components of the vector $K$.\newline

			\par In order to describe the state of a particle using the \textbf{vector space basis}, the particle's state can be written using the equation \ref{eq:ketNotationGeneric}.

			\begin{equation}
				| \psi \rangle = \alpha . | V_1 \rangle + \beta . | V_2 \rangle
				\label{eq:ketNotationGeneric}
			\end{equation}

			\par With $\alpha, \beta$ and $\gamma$ known as \textbf{probability amplitudes}.\newline

			\par Simply put, the equation \ref{eq:ketNotationGeneric} and its respective plot depicted in figure \ref{fig:diracPlot} can be interpreted as follows: \textit{The state of $\psi$ within the vector space defined by vectors $V_1$ and $V_2$ is a linear combination ($\alpha$ times vector $V_1$ plus $\beta$ times vector $V_2$)}.

			\begin{figure}[h]
				\centering
				\includegraphics{images/diracPlot}
				\caption{Visual representation of Equation \ref{eq:ketNotationGeneric}. Vectors $V_1$ and $V_2$ must be orthogonal and normalized.}
				\label{fig:diracPlot}
			\end{figure}

			\par In quantum mechanics, a state can be represented by different levels depending on the system. For instance, the previous example illustrated a representation of a \textbf{two-level state system}. However, it is important to note that a state can be extended to include one, two, or even multiple levels ($n$).\newline

			\par To represent multiple-level states, a common approach is to utilize a vector space with a basis that corresponds to the distinct levels of the system. Each level is associated with a specific vector in the vector space.\newline

			\par For instance, consider a three-level system. We can define a basis as ${|V_0\rangle, |V_1\rangle, |V_2\rangle}$, where $|V_0\rangle$, $|V_1\rangle$, and $|V_2\rangle$ represent the vectors associated with the respective levels.

			\par In general, a state in a multiple-level system can be expressed as a linear combination of these basis vectors:

			\begin{equation}
				|\psi\rangle = \alpha |V_0\rangle + \beta |V_1\rangle + \gamma |V_2\rangle \qquad,
				\label{eq:multipleLevelState}
			\end{equation}

			\par Here, $\alpha$, $\beta$, and $\gamma$ are complex coefficients that determine the \textbf{amplitudes} of the state in each level.\newline

			\par Being $N=$ the number of levels and $a \in (\alpha, \beta, \gamma, ... , \omega)$ sometimes the equation \ref{eq:multipleLevelState} can be represented as a sum:

			\begin{equation}
				|\psi\rangle = \sum_{i=1}^{N} a_i . |V_i\rangle = 
				\begin{bmatrix}
					a_{1} \\
					a_{2} \\
					\vdots \\
					a_{n}
				\end{bmatrix}
				\label{eq:multipleLevelStateSum}
			\end{equation}

			\par Some authors prefers to use $|0\rangle, |1\rangle, |2\rangle, \dots , |n\rangle / n < N$ instead of $|V_0\rangle,|V_1\rangle, \dots, |V_n\rangle / n < N$.\newline

			\par Note that this representation is nothing more than another way of defining a \textbf{vector} considering a \textbf{vector space basis} already used in linear algebra, as shown in Equation \ref{eq:linearCombination}!\newline

			\begin{equation}
				\label{eq:linearCombination}
				\begin{aligned}
					|\psi\rangle =
					\begin{bmatrix}
						a_{1} \\
						a_{2} \\
						\vdots \\
						a_{n}
					\end{bmatrix} =
					a_1 . \begin{bmatrix}
						1\\
						0\\
						\vdots \\
						0
					\end{bmatrix}+
					a_2 . \begin{bmatrix}
						0\\
						1\\
						\vdots \\
						0
					\end{bmatrix}+
					\dots + a_n . \begin{bmatrix}
						0\\
						0\\
						\vdots \\
						1
					\end{bmatrix}
				\end{aligned}
			\end{equation}

			\par Remember this basis vectors $
			\begin{bmatrix}
				1\\
				0\\
				\vdots \\
				0
			\end{bmatrix}, 
			\begin{bmatrix}
				0\\
				1\\
				\vdots \\
				0
			\end{bmatrix}, \dots,
			\begin{bmatrix}
				0\\
				0\\
				\vdots \\
				1
			\end{bmatrix}$ are the \textbf{system possible states}!

		\subsection{Probability amplitude}
			\label{sec:probAmpli}
			\par As shown in a previous section, the probability amplitudes represented by the values $[a_1,a_2,a_3, \dots, a_n]$ reside in the complex number set $\mathbb{C}$. In some cases, $a_n$ may have a value in the format $u + vi$, where $u, v \in \mathbb{R}$ and $i = \sqrt{-1}$, indicating a number that is \textbf{not real}. Therefore, in this context, it is not possible to assign a meaningful interpretation to these numbers.\newline

			\par To address this issue, we need to introduce the concept of the complex conjugate of a number. For a number $z = u + vi$, where $z \in \mathbb{C}$, its conjugate is denoted as $z^*$, where $z^* \in \mathbb{C}$, and it is given by $z^* = u - vi$. In summary, if $z = u + vi$, then $z^* = u - vi$. \newline

			\par By applying the concept of the complex conjugate, we can see that $z \cdot z^* = (u + vi) \cdot (u - vi) = u^2 + v^2 \in \mathbb{R}$. This resolves the problem, as the product $z \cdot z^*$ is a real number!\newline

			\par Finally, let \(P(S_i)\) be a function that calculates the probability of the state \(S_i\) occurring. For instance, if \(|\psi\rangle = \alpha |S_0\rangle + \beta |S_1\rangle\), then \(P(S_0)\) and \(P(S_1)\) are defined as shown in equations \ref{eq:state0} and \ref{eq:state1} \textbf{REFERÊNCIA AQUI}.

			\begin{equation}
				\label{eq:state0}
				P(S_0) = \alpha . \alpha^* 
			\end{equation}

			\begin{equation}
				\label{eq:state1}
				P(S_1) = \beta . \beta^* 
			\end{equation}

		\subsection{Bra notation}
			\label{sec:braNotation}

			\par When calculating the probability of a specific state, as discussed in section \ref{sec:probAmpli}, a new question arises: \textit{What if there is a need to calculate the total probability of all possible states?} This is where the "bra" notation comes into play!\newline

			\par As we already know, a vector of all probability amplitudes, denoted as $|\psi\rangle$, is given by equation \ref{eq:multipleLevelStateKet}. So, for these "ket" states, there is an associated "bra" notation defined in equation \ref{eq:multipleLevelStateBra} \textbf{REFERENCIA AQUI}.

			\begin{equation}
				|\psi\rangle = \sum_{i=1}^{N} a_i . |V_i\rangle = 
				\begin{bmatrix}
					a_{1} \\
					a_{2} \\
					\vdots \\
					a_{n}
				\end{bmatrix}
				\label{eq:multipleLevelStateKet}
			\end{equation}

			\begin{equation}
				\langle\psi| = \sum_{j=1}^{N} a_j^* . \langle V_j| = [a_1^*, a_2^*, \dots, a_n^*]
				\label{eq:multipleLevelStateBra}
			\end{equation}

			\par Considering the equations \ref{eq:multipleLevelStateKet} and \ref{eq:multipleLevelStateBra}, it is now possible, based on what was discussed in section \ref{sec:probAmpli}, to derive a new equation, denoted as \ref{eq:multipleLevelStateBraKet}. This sum corresponds to the overall probability of the system existing in this particular configuration!

			\begin{equation}
				\begin{aligned}
					\langle\psi|\psi\rangle = [a_1^*, a_2^*, \dots, a_n^*] . \begin{bmatrix}
						a_{1} \\
						a_{2} \\
						\vdots \\
						a_{n}
					\end{bmatrix} = &\\ 
					a_1^*.a_1 + a_2^*.a_2 + \dots + a_n^*.a_n = &\\
					P(S_1) + P(S_2) + \dots + P(S_n) = &\\
					\sum_{i}^{N} P(S_i)
				\end{aligned}
				\label{eq:multipleLevelStateBraKet}
			\end{equation}

			\par ----------------------------------------------------------------

			\par Bra notation allows for the calculation of inner products between vectors. The inner product of a bra vector $\langle \psi |$ and a ket vector $| \phi \rangle$ is denoted as $\langle \psi | \phi \rangle$. The inner product yields a complex number, providing information about the similarity or overlap between the two vectors.\newline

			\par Bra notation also enables the representation of operators in quantum mechanics. An operator can be represented by a matrix, and its action on a vector is given by the product of the operator matrix and the ket vector. For example, the action of an operator $\hat{A}$ on a ket vector $| \psi \rangle$ is represented as $\hat{A} | \psi \rangle$.\newline

			\par By following these principles, multiple-level states can be effectively represented, with the basis vectors and coefficients adjusted to accommodate the specific number of levels present in the system.\newline

	\section{Heisenberg principle}

		\par This idea can be explained in the context of analyzing variable signals. The challenge lies in determining both the frequencies and their corresponding locations within the signal.\newline

		\par To address this challenge, a tool called the Fourier transformation can be employed. Typically, when analyzing a signal, it is common practice to create a "window," which represents the segment size that will be processed. The largest window encompasses all the values of the signal, while the smallest window contains only one value (in the case of discrete sampling).\newline

		\par In Figure \ref{fig:noisysignal}, a multi-frequency signal is depicted. Applying the Fourier transformation to the entire signal enables the discovery of all the frequencies present, but it does not provide information about where these frequencies occur within the signal. On the other hand, using a small window allows us to identify the locations of certain high-frequency signal components, but it hinders our ability to determine the locations of lower frequencies, as illustrated in Figure \ref{fig:windowednoisysignal}.\newline

		\par In short, when it comes to analyzing signals, the size of the window plays a crucial role. A larger window facilitates more accurate frequency measurement but sacrifices the ability to determine their precise locations. Conversely, a smaller window provides better localization but compromises the accuracy of frequency measurements.

		\begin{figure}[h]
			\centering
			\includegraphics[width=1\linewidth]{images/noisySignal}
			\caption[Multi-frequency signal]{Multi-frequency signal. Source: \cite{olama2011design}}
			\label{fig:noisysignal}
		\end{figure}

		\begin{figure}[!h]
			\centering
			\includegraphics[width=0.1\linewidth]{images/windowedNoisySignal}
			\caption[Windowed signal]{Small windowed signal. Adapted from: \cite{olama2011design}}
			\label{fig:windowednoisysignal}
		\end{figure}

	\section{Quantum Bit (Qubit)}
		\par Qubits, unlike classical bits, do not have a defined state. They exist in a \textbf{superposition} of infinite states between 1 and 0 until they are measured, at which point their state collapses to either 1 or 0.

		\par In quantum mechanics, the state of a qubit can be represented using Dirac notation, as shown in Equation \ref{eq:ketNotation}.

		\begin{equation}
			| \psi \rangle = \alpha . | 1 \rangle + \beta . | 0 \rangle
			\label{eq:ketNotation}
		\end{equation}

		\par Simply put, Equation \ref{eq:ketNotation} can be interpreted as follows: The state of $\psi$ within the vector space defined by vectors 1 and 2 is a linear combination of $\alpha$ times vector 1 and $\beta$ times vector 0. It is important to note that these vectors must be orthogonal and normalized. A more generic way to write Equation \ref{eq:ketNotation} is shown in Equation \ref{eq:ketNotationGeneric} and visually represented in Figure \ref{fig:diracPlot}.

		\begin{equation}
			| \psi \rangle = \alpha . | V_1 \rangle + \beta . | V_2 \rangle
			\label{eq:ketNotationGeneric2}
		\end{equation}

		\begin{figure}[h]
			\centering
			\includegraphics{images/diracPlot}
			\caption{Visual representation of Equation \ref{eq:ketNotationGeneric}. Vectors $V_1$ and $V_2$ must be orthogonal and normalized. $\alpha$ and $\beta$ represent the probabilities associated with the state $\psi$.}
			\label{fig:diracPlot2}
		\end{figure}

		\par When measured, a qubit will collapse to one of the basis states, either 1 or 0, with probabilities given by the magnitudes of $\alpha$ and $\beta$. Note the word \textbf{probability}, in this scope it means that due to \textbf{Heisenberg Principle} is not possible to known exactly the position and velocity of particle at the same time!

		\subsection{Qubit capacity}
			\begin{itemize}
				\item 1 qubit = 2 bits
				\item 2 qubits = 4 bits
				\item 3 qubits = 8 bits (1 byte)
				\item 4 qubits = 16 bits
				\item 13 qubits = 8.192 bits (1kb)
				\item 23 qubits = 8.388.608 bits (1mb)
				\item 33 qubits = 8.589.934.592 bits (1gb)
				\item 43 qubits = 8.796.093.022.208 bits (1 terabyte)
				\item n qubits = $2^n$ bits
			\end{itemize}

	\section{Quantum Superposition}
		\par The superposition principle.

	\section{Quantum Entanglement}

	\section{Quantum operations}
		\label{sec:quantumOperations}

		\par The state vector can be subject to \textbf{quantum operations}, such as unitary transformations and measurements. These operations modify the state vector or extract information from it, allowing for the study of the system's properties and the prediction of measurement outcomes.

	\section{Examples}

	\section{Conclusion}
		\par Quantum mechanics provides a fascinating framework for understanding the behavior of matter and energy at the microscopic level. The concepts of superposition and entanglement, in particular, are crucial for the development of quantum computation. By leveraging these concepts, quantum computers have the potential to solve certain problems much more efficiently than classical computers. By familiarizing ourselves with the basic principles of quantum mechanics, computer scientists can effectively explore and harness the power of quantum computation.

	\bibliographystyle{plain}
	\bibliography{references.bib}
\end{document}

### 2023-06-12 documentation/article02/code/integral.cpp

@@ -1,58 +1,0 @@
-#include <cmath>
-#include <iostream>
-#include <ostream>
-
-#ifdef _OPENMP
-	#include <omp.h>
-#endif
-
-inline long double f(long double x)
-{
-    return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
-}
-
-inline long double integrate(long double (*f)(long double),
-                             double lowerLimit,
-                             double upperLimit,
-                             int partitions)
-{
-    double h = (upperLimit - lowerLimit) / partitions;
-    long double sum = 0;
-
-#pragma default(none) shared(soma)
-    {
-#pragma omp parallel
-        {
-#pragma omp for reduction(+ : sum)
-            for (int k = 1; k < partitions - 1; k++) {
-                sum += f(lowerLimit + k * h);
-            }
-        }
-#pragma omp barrier
-
-        sum *= 2;
-        sum += (f(lowerLimit) + f(upperLimit));
-        return (h / 2) * sum;
-    }
-}
-
-int main()
-{
-    double lowerLimit = 0, upperLimit = 100;
-
-    int partitions = 0;
-
-    partitions = upperLimit / 0.000001;
-    std::cout << "Interval: " << 0.000001 << " Partitions: " << partitions
-              << " - Result: " << integrate(f, lowerLimit, upperLimit, partitions) << std::endl;
-
-    partitions = upperLimit / 0.00001;
-    std::cout << "Interval: " << 0.00001 << " Partitions: " << partitions
-              << " - Result: " << integrate(f, lowerLimit, upperLimit, partitions) << std::endl;
-
-    partitions = upperLimit / 0.0001;
-    std::cout << "Interval: " << 0.0001 << " Partitions: " << partitions
-              << " - Result: " << integrate(f, lowerLimit, upperLimit, partitions) << std::endl;
-
-    return 0;
-}

### 2023-06-17 documentation/article02/presentation/apresentacao.tex
Almost ending, presentation started

\input{preamble.tex}

\title{Introdução à Computação Quântica}

\institute{Universidade Estadual Paulista Júlio de Mesquita Filho}

\author{André Furlan - ensismoebius@gmail.com}

\logo{\includegraphics[height=1cm]{unesp.png}}
\date{\the\year}

\begin{document}

	\frame{\titlepage}

	\begin{frame}{Introdução}
		\begin{itemize}
			\item A mecânica quântica surgiu no século XX, juntamente com os computadores clássicos.
			\item A computação quântica tem recebido atenção significativa devido ao seu potencial para resolver problemas complexos de forma mais eficiente do que os computadores clássicos.
			\item A computação quântica utiliza os princípios da mecânica quântica para realizar cálculos usando bits quânticos ou qubits.
		\end{itemize}
	\end{frame}

	\begin{frame}{Estados Quânticos}
		\begin{itemize}
			\item Os estados quânticos são representações matemáticas que fornecem informações sobre as propriedades de um sistema físico, como a posição ou o momento de uma partícula.
			\item Os estados quânticos são descritos usando vetores de estado, que são frequentemente representados usando a notação de Dirac "ket" (por exemplo, $|\psi\rangle$).
			\item Ao contrário dos bits clássicos, que podem representar apenas 0 ou 1, os estados quânticos podem existir em uma superposição, onde eles representam simultaneamente vários estados possíveis.
		\end{itemize}
	\end{frame}

	\begin{frame}{Espaço Vetorial Complexo}
		\begin{itemize}
			\item Os estados quânticos são representados usando espaços vetoriais complexos.
			\item Um espaço vetorial complexo é uma estrutura matemática que segue regras específicas, incluindo comutatividade e associatividade, permitindo a combinação de vetores.
			\item Os vetores de base dentro do espaço vetorial representam os estados possíveis do sistema quântico, e as combinações lineares desses vetores de base representam estados quânticos arbitrários.
		\end{itemize}
	\end{frame}

	\begin{frame}{Notação de Dirac}
		\begin{itemize}
			\item A notação de "BraKet" de Dirac, desenvolvida pelo físico Paul Dirac, é uma notação comum usada na mecânica quântica.
			\item A notação "ket" ($|\psi\rangle$) representa o estado de uma partícula dentro do espaço vetorial.
			\item A notação "bra" ($\langle\psi|$) representa o transposto conjugado do vetor ket.
			\item O produto interno de um bra e um ket fornece a amplitude de probabilidade para um determinado estado.
		\end{itemize}
	\end{frame}

	\begin{frame}{Amplitude de Probabilidade}
		\begin{itemize}
			\item As amplitudes de probabilidade são números complexos que descrevem a probabilidade de ocorrência de um estado quântico específico.
			\item As amplitudes de probabilidade são calculadas usando o produto interno de um bra e um ket.
			\item As amplitudes de probabilidade são normalizadas para garantir que a soma das probabilidades de todos os estados possíveis seja igual a 1.
		\end{itemize}
	\end{frame}

	\begin{frame}{Operadores Lineares}
		\begin{itemize}
			\item Os operadores lineares são ferramentas matemáticas usadas para descrever transformações em espaços vetoriais.
			\item Na mecânica quântica, os operadores lineares representam observáveis físicos, como posição, momento e energia.
			\item Os operadores lineares podem ser aplicados a vetores de estado para obter novos vetores de estado resultantes da transformação.
		\end{itemize}
	\end{frame}

	\begin{frame}{Operadores Lineares (cont.)}
		\begin{itemize}
			\item Exemplos de operadores lineares comumente usados na mecânica quântica incluem o operador de posição, operador de momento e operador de energia.
			\item Esses operadores têm propriedades específicas, como hermiticidade e comutatividade, que são importantes na mecânica quântica.
		\end{itemize}
	\end{frame}

	\begin{frame}{Portas Quânticas}
		\begin{itemize}
			\item Portas quânticas são análogas aos circuitos lógicos em computadores clássicos.
			\item Elas são operadores lineares que agem em qubits individuais ou em múltiplos qubits.
			\item As portas quânticas permitem a transformação dos estados quânticos, possibilitando várias operações computacionais.
			\item Algumas portas quânticas conhecidas incluem a porta Hadamard, porta Pauli-X e a porta CNOT.
		\end{itemize}
	\end{frame}

	\begin{frame}{Portas Quânticas (cont.)}
		\begin{itemize}
			\item Porta Hadamard: Ela cria uma superposição de estados em um qubit. É representada pela seguinte matriz:
			\[
			H = \frac{1}{\sqrt{2}}\begin{bmatrix}
				1 & 1 \\
				1 & -1 \\
			\end{bmatrix}
			\]
			\item Porta Pauli-X: Conhecida como porta NOT quântica, ela inverte o estado de um qubit. É representada pela seguinte matriz:
			\[
			X = \begin{bmatrix}
				0 & 1 \\
				1 & 0 \\
			\end{bmatrix}
			\]
			\item Porta CNOT: A porta Controlled NOT opera em dois qubits, onde um atua como controle e o outro como alvo. É representada pela seguinte matriz:
			\[
			CNOT = \begin{bmatrix}
				1 & 0 & 0 & 0 \\
				0 & 1 & 0 & 0 \\
				0 & 0 & 0 & 1 \\
				0 & 0 & 1 & 0 \\
			\end{bmatrix}
			\]
		\end{itemize}
	\end{frame}

	\begin{frame}{Conclusão}
		\begin{itemize}
			\item A computação quântica é um campo promissor que utiliza os princípios da mecânica quântica para realizar cálculos mais eficientes.
			\item Os estados quânticos, espaços vetoriais complexos, operadores lineares e portas quânticas são conceitos fundamentais na computação quântica.
			\item As portas quânticas permitem a realização de operações computacionais, e exemplos conhecidos incluem a porta Hadamard, porta Pauli-X e porta CNOT.
			\item A compreensão desses conceitos é essencial para explorar o potencial da computação quântica e suas aplicações futuras.
		\end{itemize}
	\end{frame}

	\begin{frame}[allowframebreaks]
		\frametitle{Referências}
		\bibliography{bibliography.bib}
	\end{frame}

\end{document}

### 2023-06-17 documentation/article02/text/article.tex
Almost ending, presentation started

--- 2023-06-12 documentation/article02/article.tex
@@
 This document aims to provide a brief review of these concepts to enable those interested in quantum computation to understand and utilize its ideas effectively.
+For a broader explanation of quantum computing, I recommend the insightful work "Introdução à Computação Quântica" by Wagner Jorcuvich Nunes da Silva (2018) \cite{da2018introduccao}.
 	\end{abstract}
@@
 This concept that will be discussed deeply in section \ref{sec:probAmpli} is known as \textbf{probability amplitude} and it employs two notations: the "ket" and the "bra".
-These notations are commonly referred to as the \textbf{Dirac's "BraKet" notation}.
+These notations are commonly referred to as the \textbf{Dirac's "BraKet" notation} \cite{notacaoDirac}.
 
@@
 			\begin{equation}
-				|K\rangle = \vec{K} = \{a, b, c, d\}
+				|K\rangle = \vec{K} = \begin{bmatrix}
+					a \\
+					b \\
+					c \\
+					d
+				\end{bmatrix}
 				\label{eq:ketNotationVector}
@@
 
-			\par Where $a, b, c, d \in \mathbb{Q}$ and are components of the vector $K$.\newline
+			\par Where $a, b, c, d \in \mathbb{Q}$ and are components of the \textbf{column vector} $K$.\newline
 
@@
 			\par Finally, let \(P(S_i)\) be a function that calculates the probability of the state \(S_i\) occurring.
-For instance, if \(|\psi\rangle = \alpha |S_0\rangle + \beta |S_1\rangle\), then \(P(S_0)\) and \(P(S_1)\) are defined as shown in equations \ref{eq:state0} and \ref{eq:state1} \textbf{REFERÊNCIA AQUI}.
+For instance, if \(|\psi\rangle = \alpha |S_0\rangle + \beta |S_1\rangle\), then \(P(S_0)\) and \(P(S_1)\) are defined as shown in equations \ref{eq:state0} and \ref{eq:state1} \cite{binney2013physics}.
 
@@
 
+			\par Briefly, given a ket $|\psi\rangle$ (as shown in equation \ref{eq:ket}), its corresponding bra is denoted by the conjugate transpose of $|\psi\rangle$ (equation \ref{eq:bra}).
+
+			\begin{equation}
+				\label{eq:ket}
+				|\psi\rangle = 
+				\begin{bmatrix}
+					u + vo \\
+					k + ji \\
+					o + li \\
+					d + mi
+				\end{bmatrix} \in \mathbb{C}
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:bra}
+				\langle\psi| = [u - vo, k - ji, o - li, d - mi]
+			\end{equation}
+
 			\par When calculating the probability of a specific state, as discussed in section \ref{sec:probAmpli}, a new question arises: \textit{What if there is a need to calculate the total probability of all possible states?} This is where the "bra" notation comes into play!\newline
@@
 			\par As we already know, a vector of all probability amplitudes, denoted as $|\psi\rangle$, is given by equation \ref{eq:multipleLevelStateKet}.
-So, for these "ket" states, there is an associated "bra" notation defined in equation \ref{eq:multipleLevelStateBra} \textbf{REFERENCIA AQUI}.
+So, for these "ket" states, there is an associated "bra" notation defined in equation \ref{eq:multipleLevelStateBra} \cite{notacaoDirac}
 
@@
 					P(S_1) + P(S_2) + \dots + P(S_n) = &\\
-					\sum_{i}^{N} P(S_i)
+					\sum_{i=1}^{N} P(S_i)
 				\end{aligned}
@@
 			\end{equation}
-
-			\par ----------------------------------------------------------------
 
@@
 
-			\par Bra notation also enables the representation of operators in quantum mechanics.
-An operator can be represented by a matrix, and its action on a vector is given by the product of the operator matrix and the ket vector.
-For example, the action of an operator $\hat{A}$ on a ket vector $| \psi \rangle$ is represented as $\hat{A} | \psi \rangle$.\newline
+			\par For instance, given a vector 
+				$|\alpha\rangle= 
+				\begin{bmatrix}
+					1 \\
+					0
+				\end{bmatrix}
+				$
+				 and 
+				 $|\beta\rangle = \begin{bmatrix}
+					0\\
+					1
+				\end{bmatrix}$ 
+				using the "BraKet" operators to represent the inner product, it will end up like in equation \ref{eq:braket}.
 
-			\par By following these principles, multiple-level states can be effectively represented, with the basis vectors and coefficients adjusted to accommodate the specific number of levels present in the system.\newline
+			\begin{equation}
+				\label{eq:braket}
+				\langle \alpha | \beta \rangle = 
+				[1, 0] .
+				\begin{bmatrix}
+					0 \\
+					1
+				\end{bmatrix} = 1.0 + 0.1 = 0
+			\end{equation}
 
@@
 	\section{Quantum Bit (Qubit)}
-		\par Qubits, unlike classical bits, do not have a defined state.
-They exist in a \textbf{superposition} of infinite states between 1 and 0 until they are measured, at which point their state collapses to either 1 or 0.
-
-		\par In quantum mechanics, the state of a qubit can be represented using Dirac notation, as shown in Equation \ref{eq:ketNotation}.
+		\par Qubits, unlike classical bits, do not have a defined state and exist in a superposition of states 1 and 0 until they are measured, at which point their state collapses to either 1 or 0 with probabilities given by $|\alpha|^2$ and $|\beta|^2$ respectively, as represented in Equation \ref{eq:qubit} and Figure \ref{fig:qubitplot} \cite{da2018introduccao}.
 
 		\begin{equation}
-			| \psi \rangle = \alpha . | 1 \rangle + \beta . | 0 \rangle
-			\label{eq:ketNotation}
+			| \psi \rangle = \alpha . | 0 \rangle + \beta . | 1 \rangle = 
+			\alpha . \begin{bmatrix}
+				1 \\
+				0
+			\end{bmatrix} + \beta .	\begin{bmatrix}
+				0 \\
+				1
+			\end{bmatrix} = 
+			\begin{bmatrix}
+				\alpha \\
+				\beta
+			\end{bmatrix}
+			\label{eq:qubit}
 		\end{equation}
 
-		\par Simply put, Equation \ref{eq:ketNotation} can be interpreted as follows: The state of $\psi$ within the vector space defined by vectors 1 and 2 is a linear combination of $\alpha$ times vector 1 and $\beta$ times vector 0.
-It is important to note that these vectors must be orthogonal and normalized.
-A more generic way to write Equation \ref{eq:ketNotation} is shown in Equation \ref{eq:ketNotationGeneric} and visually represented in Figure \ref{fig:diracPlot}.
-
-		\begin{equation}
-			| \psi \rangle = \alpha . | V_1 \rangle + \beta . | V_2 \rangle
-			\label{eq:ketNotationGeneric2}
-		\end{equation}
+		\par The coefficients $\alpha$ and $\beta$ are complex numbers such that ${|\alpha|}^2 + {|\beta|}^2 = 1$, the vector 
+		$\begin{bmatrix}
+			\alpha \\
+			\beta 
+		\end{bmatrix}$ is the qubit itself and, by convention, $| 1 \rangle = 
+		\begin{bmatrix}
+			0 \\
+			1
+		\end{bmatrix} = 1$ and $| 0 \rangle = 
+		\begin{bmatrix}
+			1 \\
+			0
+		\end{bmatrix} = 0$
 
@@
 			\centering
-			\includegraphics{images/diracPlot}
-			\caption{Visual representation of Equation \ref{eq:ketNotationGeneric}.
-Vectors $V_1$ and $V_2$ must be orthogonal and normalized.
-$\alpha$ and $\beta$ represent the probabilities associated with the state $\psi$.}
-			\label{fig:diracPlot2}
+			\includegraphics{images/qubitPlot}
+			\caption{Visual representation of Equation \ref{eq:qubit}. Vectors $1$ and $2$ must be orthogonal and normalized. $\alpha$ and $\beta$ represent the probabilities associated with the state $\psi$.}
+			\label{fig:qubitplot}
 		\end{figure}
 
-		\par When measured, a qubit will collapse to one of the basis states, either 1 or 0, with probabilities given by the magnitudes of $\alpha$ and $\beta$.
-Note the word \textbf{probability}, in this scope it means that due to \textbf{Heisenberg Principle} is not possible to known exactly the position and velocity of particle at the same time!
+		\par The Equation \ref{eq:qubit} can be interpreted as follows: The \textbf{non-measured} state of $\psi$ within the vector space defined by vectors $| 1 \rangle$ and $| 0 \rangle$ is a linear combination of them  ($\alpha$ times vector $| 0 \rangle$ plus $\beta$ times vector $| 1 \rangle$).\newline
 
-		\subsection{Qubit capacity}
+		\par It's important to note that in Figure \ref{fig:qubitplot} and Equation \ref{eq:qubit}, a quantum state is often in a superposition because $\alpha$ and $\beta$ can take on multiple values along the bases axis.
+For example, if $\alpha = 1$ and $\beta = 0$, the qubit has a $100\%$ probability of being a 0 and $0\%$ probability of being a 1.
+Conversely, if $\alpha = \frac{1}{\sqrt{2}}$ and $\beta = \frac{1}{\sqrt{2}}$, there is a $50\%$ probability of being 1 and a $50\%$ of being 0.
+		\newline
+
+		\par Qubits have the ability to store and process information differently than classical bits due to superposition.
+Their capacity, in terms of information storage, is significantly larger compared to classical bits.
+However, there is a catch: when qubits are measured, they collapse into classical bits.
+Therefore, currently, it is not possible to take full advantage of the quantum properties for data storage purposes.
+Furthermore, qubits can store non-binary data, such as $\begin{bmatrix} \dfrac{1}{\sqrt{2}} \\\\ \dfrac{1}{\sqrt{2}} \end{bmatrix}$ or $\begin{bmatrix} \dfrac{1}{2} \\\\ \dfrac{\sqrt{3}}{2} \end{bmatrix}$.\newline
+
+		\par Another challenge that arises with systems containing many qubits is decoherence.
+Due to the limitations of current technology, as the number of qubits increases, so does the presence of noise.
+This leads to a higher likelihood of random collapses and decoherence between the states of qubits.\newline
+
+		\par The List bellow shows the qubits capacity.
+
 		\begin{itemize}
+			\label{lst:qubits}
 			\item 1 qubit = 2 bits
@@
 			\item 4 qubits = 16 bits
-				\item 13 qubits = 8.192 bits (1kb)
-				\item 23 qubits = 8.388.608 bits (1mb)
-				\item 33 qubits = 8.589.934.592 bits (1gb)
-				\item 43 qubits = 8.796.093.022.208 bits (1 terabyte)
+			\item 13 qubits = 8,192 bits (1 kilobyte)
+			\item 23 qubits = 8,388,608 bits (1 megabyte)
+			\item 33 qubits = 8,589,934,592 bits (1 gigabyte)
+			\item 43 qubits = 8,796,093,022,208 bits (1 terabyte)
 			\item n qubits = $2^n$ bits
 		\end{itemize}
-
-	\section{Quantum Superposition}
-		\par The superposition principle.
-
-	\section{Quantum Entanglement}
 
@@
 
-		\par The state vector can be subject to \textbf{quantum operations}, such as unitary transformations and measurements.
-These operations modify the state vector or extract information from it, allowing for the study of the system's properties and the prediction of measurement outcomes.
+		\par The state vector of a quantum system can undergo various quantum operations, including unitary transformations and measurements.
+These operations enable the modification of the state vector and the extraction of information from it, facilitating the study of system properties and the prediction of measurement outcomes.\newline
 
-	\section{Examples}
+		\par Quantum operations are implemented using quantum gates, which are represented by unitary matrices.
+These gates serve as unitary linear transformations that preserve orthogonality, ensuring that the quantum state remains normalized and the probabilities of different outcomes are preserved.\newline
+
+		\par To preserve the normalization a property must be observed: Being $M$ a matrix and $M^*$ its conjugate transpose and $I$ an indentity matrix, a matrix $M$ is unitary if $M.M^* = I$ \cite{da2018introduccao}.
+
+		\subsection{Quantum gates}
+
+			\par As mentioned earlier, quantum gates are represented as matrices, and their number is infinite because the values of the matrix cells can vary across the real numbers ($\mathbb{R}$).
+Although there is a wide variety of quantum gates and considering $| \psi \rangle = \alpha .
+| 0 \rangle + \beta .
+| 1 \rangle$, here are a few examples.\newline
+
+			\par The \textbf{identity gate} in Equation \ref{eq:indentity} alters the qubit state as shown in Equation \ref{eq:indentityApplyed}.
+			\begin{equation}
+				\label{eq:indentity}
+					I : 
+					\begin{bmatrix}
+						1& 0 \\
+						0& 1
+					\end{bmatrix}
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:indentityApplyed}
+				I|\psi\rangle = \begin{bmatrix}
+					1& 0 \\
+					0& 1
+				\end{bmatrix} . \begin{bmatrix}
+					\alpha \\
+					\beta
+				\end{bmatrix} = \begin{bmatrix}
+					\alpha \\
+					\beta
+				\end{bmatrix} = |\psi\rangle
+			\end{equation}
+
+			\par The \textbf{flip gate} in Equation \ref{eq:flip} flips the qubit state as shown in Equation \ref{eq:indentityApplyed}.
+			\begin{equation}
+				\label{eq:flip}
+				X : 
+				\begin{bmatrix}
+					0& 0 \\
+					1& 0
+				\end{bmatrix}
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:flipApplyed}
+				X|\psi\rangle = \begin{bmatrix}
+					0& 1 \\
+					1& 0
+				\end{bmatrix} . \begin{bmatrix}
+					\alpha \\
+					\beta
+				\end{bmatrix} = \begin{bmatrix}
+					\beta \\
+					\alpha
+				\end{bmatrix}
+			\end{equation}
+
+			\par The \textbf{Hadamard gate} in Equation \ref{eq:hadamard} puts the qubit state in to \textbf{superposition} \cite{qcfcs}  with the same probability for the two states \cite{da2018introduccao} as shown in Equation \ref{eq:hadamardApplyed}.
+			\begin{equation}
+				\label{eq:hadamard}
+				H : \dfrac{1}{\sqrt{2}}.
+				\begin{bmatrix}
+					1& 1 \\
+					1& -1
+				\end{bmatrix}
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:hadamardApplyed}
+				\begin{aligned}
+					&H|\psi\rangle = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
+						1& 1 \\
+						1& -1
+					\end{bmatrix} . \begin{bmatrix}
+						\alpha \\
+						\beta
+					\end{bmatrix} = \\
+					&\begin{bmatrix}
+						\dfrac{1}{\sqrt{2}} &\dfrac{1}{\sqrt{2}} \\
+						\dfrac{1}{\sqrt{2}} &\dfrac{-1}{\sqrt{2}}
+					\end{bmatrix} .
+					\begin{bmatrix}
+						\alpha \\
+						\beta
+					\end{bmatrix} = 
+					\begin{bmatrix}
+						\alpha . \dfrac{1}{\sqrt{2}} + \beta.\dfrac{1}{\sqrt{2}} \\
+						\alpha . \dfrac{1}{\sqrt{2}} + \beta.\dfrac{-1}{\sqrt{2}}
+					\end{bmatrix} = \\
+					&\begin{bmatrix}
+						\dfrac{1}{\sqrt{2}} . (\alpha + \beta) \\
+						\dfrac{1}{\sqrt{2}} . (\alpha - \beta)
+					\end{bmatrix}
+				\end{aligned}
+			\end{equation}
+
+			\par To further illustrate this, let's consider a numeric example.
+We start with a qubit state $\begin{bmatrix} 1 \\ 0 \end{bmatrix}$, which is not in superposition.
+We can apply the Hadamard gate to put it into superposition.
+The resulting state is given by Equation \ref{eq:hadamardApplyedNum}.
+
+			\begin{equation}
+				\label{eq:hadamardApplyedNum}
+				\begin{aligned}
+					&H|\psi\rangle = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
+						1& 1 \\
+						1& -1
+					\end{bmatrix} . \begin{bmatrix}
+						1 \\
+						0
+					\end{bmatrix} = \\
+					&\begin{bmatrix}
+						\dfrac{1}{\sqrt{2}} &\dfrac{1}{\sqrt{2}} \\
+						\dfrac{1}{\sqrt{2}} &\dfrac{-1}{\sqrt{2}}
+					\end{bmatrix} .
+					\begin{bmatrix}
+						1 \\
+						0
+					\end{bmatrix} = 
+					\begin{bmatrix}
+						1 . \dfrac{1}{\sqrt{2}} + 0.\dfrac{1}{\sqrt{2}} \\
+						1 . \dfrac{1}{\sqrt{2}} + 0.\dfrac{-1}{\sqrt{2}}
+					\end{bmatrix} = \\
+					&\begin{bmatrix}
+						\dfrac{1}{\sqrt{2}} . (1 + 0) \\
+						\dfrac{1}{\sqrt{2}} . (1 - 0)
+					\end{bmatrix} = 
+					\begin{bmatrix}
+						\dfrac{1}{\sqrt{2}} \\
+						\dfrac{1}{\sqrt{2}}
+					\end{bmatrix}
+				\end{aligned}
+			\end{equation}
+
+			\par The Hadamard gate possesses another interesting property: when applied to a quantum bit, it can transform it from a superposition state to a classical bit with measurement, i.e.
+not triggering the collapse.
+This property makes the Hadamard gate useful for extracting results from quantum algorithms.
+Furthermore, the Hadamard gate can also perform the opposite operation: it can convert a classical bit into a superposition state.
+This allows for computation in the quantum realm, and by applying the Hadamard gate again, the results can be extracted from the superposition state \cite{qcfcs}.\newline
+
+			\par The \textbf{rotation gate} in Equation \ref{eq:rot} rotates the qubit state $\pi$ radians as shown in Equation \ref{eq:rotApplyed}.
+			\begin{equation}
+				\label{eq:rot}
+				Z : 
+				\begin{bmatrix}
+					1& 0 \\
+					0& -1
+				\end{bmatrix}
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:rotApplyed}
+				Z|\psi\rangle = \begin{bmatrix}
+					1& 0 \\
+					0& -1
+				\end{bmatrix} . \begin{bmatrix}
+					\alpha \\
+					\beta
+				\end{bmatrix} = \begin{bmatrix}
+					\alpha \\
+					-\beta
+				\end{bmatrix}
+			\end{equation}
+
+			\par The \textbf{CNOT gate} in Equation \ref{eq:cnot} creates an \textbf{entanglement} between two qubits as shown in Equation \ref{eq:cnotApplyed}.
+			\begin{equation}
+				\label{eq:cnot}
+				CNOT : 
+				\begin{bmatrix}
+					1& 0& 0& 0 \\
+					0& 1& 0& 0 \\
+					0& 0& 0& 1 \\
+					0& 0& 1& 0
+				\end{bmatrix}
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:cnotApplyed}
+				Z|\psi\rangle = \begin{bmatrix}
+					1& 0 \\
+					0& -1
+				\end{bmatrix} . \begin{bmatrix}
+					\alpha \\
+					\beta
+				\end{bmatrix} = \begin{bmatrix}
+					\alpha \\
+					-\beta
+				\end{bmatrix}
+			\end{equation}
+
+			\par Please note that, as aforementioned, this is just a sample of the possible gates, and that these ones were selected for their relevance according to the judgment of the author.
 
 	\section{Conclusion}
-		\par Quantum mechanics provides a fascinating framework for understanding the behavior of matter and energy at the microscopic level.
-The concepts of superposition and entanglement, in particular, are crucial for the development of quantum computation.
-By leveraging these concepts, quantum computers have the potential to solve certain problems much more efficiently than classical computers.
-By familiarizing ourselves with the basic principles of quantum mechanics, computer scientists can effectively explore and harness the power of quantum computation.
+		\par It is worth noting that all the examples presented in this document intentionally avoided the use of numbers from the complex set $\mathbb{Q}$.
+This choice was made to facilitate the understanding of the topic.
+However, if you are interested in delving deeper into the subject, I recommend consulting the provided references and keywords for further reading.
 
+		\par This said: That's all folks!	
 	\bibliographystyle{plain}

### 2023-06-18 documentation/article02/text/article.tex
New qubits representation added

@@
 For example, if $\alpha = 1$ and $\beta = 0$, the qubit has a $100\%$ probability of being a 0 and $0\%$ probability of being a 1.
-Conversely, if $\alpha = \frac{1}{\sqrt{2}}$ and $\beta = \frac{1}{\sqrt{2}}$, there is a $50\%$ probability of being 1 and a $50\%$ of being 0.
-		\newline
+Conversely, if $\alpha = \frac{1}{\sqrt{2}}$ and $\beta = \frac{1}{\sqrt{2}}$, there is a $50\%$ probability of being 1 and a $50\%$ of being 0.\newline
+
+		\par One good example of what a qubit can be is a photon.
+This particle has electric and magnetic components, forming an electric-magnetic field.
+We can interpret, for instance, the "vertical" component as the state 0 and the "horizontal" component as the state 1, as depicted in Figure \ref{fig:photonstate01}.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=0.7\linewidth]{images/photonState01}
+			\caption[Qubit state]{Qubit state: $\alpha$ is the probability of this qubit collapsing to 0, and $\beta$ is the probability of this qubit collapsing to 1.}
+			\label{fig:photonstate01}
+		\end{figure}
+
+		\par Note that in Figure \ref{fig:photonstate01}, there is a \textbf{superposition} of states, meaning that both 0 and 1 values can be assumed when the qubit is measured.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=0.7\linewidth]{images/photonState02}
+			\caption[Qubit state 2]{Qubit state: $\alpha$ is the probability of this qubit collapsing to 0, and $\beta$ is the probability of this qubit collapsing to 1.}
+			\label{fig:photonstate02}
+		\end{figure}
+
+		\par Note that in Figure \ref{fig:photonstate02}, there is also a \textbf{superposition}, but here the qubit is more likely to collapse to 0 because of a larger $\alpha$.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=0.7\linewidth]{images/photonState03}
+			\caption[Qubit state 3]{Qubit state: $\alpha$ is the probability of this qubit collapsing to 0, and $\beta$ is the probability of this qubit collapsing to 1.}
+			\label{fig:photonstate03}
+		\end{figure}
+
+		\par Again in Figure \ref{fig:photonstate03}, there is a \textbf{superposition}, but note that here the qubit is more likely to collapse to 1 because of a larger $\beta$.\newline
 
@@
 
-			\par The \textbf{CNOT gate} in Equation \ref{eq:cnot} creates an \textbf{entanglement} between two qubits as shown in Equation \ref{eq:cnotApplyed}.
+			\par The \textbf{conditional not gate}(CNOT) in Equation \ref{eq:cnot} creates an \textbf{entanglement} between two qubits as shown in Equation \ref{eq:cnotApplyed}.
 			\begin{equation}

### 2023-06-18 documentation/article02/text/article.tex
Article is done (for now)

@@
 
-		\begin{figure}[h]
+		\begin{figure}[H]
 			\centering
-			\includegraphics[width=0.7\linewidth]{images/photonState01}
+			\includegraphics[width=0.5\linewidth]{images/photonState01}
 			\caption[Qubit state]{Qubit state: $\alpha$ is the probability of this qubit collapsing to 0, and $\beta$ is the probability of this qubit collapsing to 1.}
@@
 
-		\par Note that in Figure \ref{fig:photonstate01}, there is a \textbf{superposition} of states, meaning that both 0 and 1 values can be assumed when the qubit is measured.
+		\par Note that in Figure \ref{fig:photonstate01}, there is a \textbf{superposition} of states, meaning that both 0 and 1 values can be assumed when the qubit is measured.\newline
 
-		\begin{figure}[h]
+		\begin{figure}[H]
 			\centering
-			\includegraphics[width=0.7\linewidth]{images/photonState02}
+			\includegraphics[width=0.5\linewidth]{images/photonState02}
 			\caption[Qubit state 2]{Qubit state: $\alpha$ is the probability of this qubit collapsing to 0, and $\beta$ is the probability of this qubit collapsing to 1.}
@@
 
-		\par Note that in Figure \ref{fig:photonstate02}, there is also a \textbf{superposition}, but here the qubit is more likely to collapse to 0 because of a larger $\alpha$.
+		\par Note that in Figure \ref{fig:photonstate02}, there is also a \textbf{superposition}, but here the qubit state is more likely to collapse to 0 because of a larger $\alpha$.\newline
 
-		\begin{figure}[h]
+		\begin{figure}[H]
 			\centering
-			\includegraphics[width=0.7\linewidth]{images/photonState03}
+			\includegraphics[width=0.5\linewidth]{images/photonState03}
 			\caption[Qubit state 3]{Qubit state: $\alpha$ is the probability of this qubit collapsing to 0, and $\beta$ is the probability of this qubit collapsing to 1.}
@@
 
-		\par Again in Figure \ref{fig:photonstate03}, there is a \textbf{superposition}, but note that here the qubit is more likely to collapse to 1 because of a larger $\beta$.\newline
+		\par Again in Figure \ref{fig:photonstate03}, there is a \textbf{superposition}, but note that here the qubit state is more likely to collapse to 1 because of a larger $\beta$.\newline
 
@@
 
-		\par Quantum operations are implemented using quantum gates, which are represented by unitary matrices.
-These gates serve as unitary linear transformations that preserve orthogonality, ensuring that the quantum state remains normalized and the probabilities of different outcomes are preserved.\newline
+		\par The Quantum operations gates serve as unitary linear transformations that preserve orthogonality, ensuring that the quantum state remains normalized and the probabilities of different outcomes are preserved.\newline
 
-		\par To preserve the normalization a property must be observed: Being $M$ a matrix and $M^*$ its conjugate transpose and $I$ an indentity matrix, a matrix $M$ is unitary if $M.M^* = I$ \cite{da2018introduccao}.
+		\par To preserve normalization, quantum gates \textbf{must} satisfy certain conditions.
+Let $G$ be a matrix representing a quantum gate and $G^*$ its conjugate transpose, and let $I$ be an identity matrix.
+A gate $G$ must be unitary being $G \cdot G^* = I$ \cite{da2018introduccao}.
 
@@
 
-			\par The \textbf{Hadamard gate} in Equation \ref{eq:hadamard} puts the qubit state in to \textbf{superposition} \cite{qcfcs}  with the same probability for the two states \cite{da2018introduccao} as shown in Equation \ref{eq:hadamardApplyed}.
+			\par The \textbf{Hadamard gate} in Equation \ref{eq:hadamard} puts the qubit state in to \textbf{superposition} \cite{qcfcs}  with the same probability for the two states \cite{da2018introduccao} allowing as shown in Equation \ref{eq:hadamardApplyed}.
 			\begin{equation}
@@
 
-			\par The \textbf{conditional not gate}(CNOT) in Equation \ref{eq:cnot} creates an \textbf{entanglement} between two qubits as shown in Equation \ref{eq:cnotApplyed}.
+			\par The \textbf{conditional not gate}(CNOT) in Equation \ref{eq:cnot} may create an \textbf{entanglement} between two qubits \cite{nielsen2002quantum}.
 			\begin{equation}
@@
 
+			\par This gate operates on \textbf{two qubits}, where one is the \textbf{control qubit} and the other is the \textbf{target qubit}.
+To represent the state of the combined qubits, a variation of ket notation is used, where two vectors are represented together.
+This is necessary because the qubits can no longer be written separately due to the entanglement this gate may cause.
+The ket variations are shown in Equations \ref{eq:ketForTwoVectors00}, \ref{eq:ketForTwoVectors01} and \ref{eq:ketForTwoVectors02}.
+
+			\begin{equation}
+				\label{eq:ketForTwoVectors00}
+				\ket{1}, \ket{0} \implies
+				\ket{10} = \begin{bmatrix}
+					0 \\
+					1 \\
+					1 \\
+					0
+				\end{bmatrix}				
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:ketForTwoVectors01}
+				\ket{0}, \ket{0} \implies
+				\ket{00} = \begin{bmatrix}
+					1 \\
+					0 \\
+					1 \\
+					0
+				\end{bmatrix}				
+			\end{equation}
+
+			\begin{equation}
+				\label{eq:ketForTwoVectors02}
+				\ket{1}, \ket{1} \implies
+				\ket{11} = \begin{bmatrix}
+					0 \\
+					1 \\
+					0 \\
+					1
+				\end{bmatrix}				
+			\end{equation}
+
+			\par The CNOT gate operates on two qubits, where the first qubit serves as the control and the second qubit as the target.
+As shown in the Equation \ref{eq:cnotApplyed}, when the control qubit is in the state $\ket{1}$, the target qubit undergoes a flip.
+However, when the control qubit is in the state $\ket{0}$, the target qubit remains unchanged.
+
 			\begin{equation}
 				\label{eq:cnotApplyed}
-				Z|\psi\rangle = \begin{bmatrix}
-					1& 0 \\
-					0& -1
-				\end{bmatrix} . \begin{bmatrix}
-					\alpha \\
-					\beta
-				\end{bmatrix} = \begin{bmatrix}
-					\alpha \\
-					-\beta
-				\end{bmatrix}
+				\begin{aligned}
+					&CNOT\ket{00} = \ket{00}, \\
+					&CNOT\ket{01} = \ket{01}, \\
+					&CNOT\ket{10} = \ket{11}, \\
+					&CNOT\ket{11} = \ket{10}
+				\end{aligned}
 			\end{equation}

### 2023-06-18 documentation/article02/text/article.tex
Ket notation fix

@@
 			\begin{equation}
-				|K\rangle = \vec{K} = \begin{bmatrix}
+				\ket{K} = \vec{K} = \begin{bmatrix}
 					a \\
@@
 			\begin{equation}
-				| \psi \rangle = \alpha . | V_1 \rangle + \beta . | V_2 \rangle
+				\ket{\psi} = \alpha . \ket{V_1} + \beta . \ket{V_2}
 				\label{eq:ketNotationGeneric}
@@
 			\par For instance, consider a three-level system.
-We can define a basis as ${|V_0\rangle, |V_1\rangle, |V_2\rangle}$, where $|V_0\rangle$, $|V_1\rangle$, and $|V_2\rangle$ represent the vectors associated with the respective levels.
+We can define a basis as ${\ket{V_0}, \ket{V_1}, \ket{V_2}}$, where $\ket{V_0}$, $\ket{V_1}$, and $\ket{V_2}$ represent the vectors associated with the respective levels.
 
@@
 			\begin{equation}
-				|\psi\rangle = \alpha |V_0\rangle + \beta |V_1\rangle + \gamma |V_2\rangle \qquad,
+				\ket{\psi} = \alpha \ket{V_0} + \beta \ket{V_1} + \gamma \ket{V_2} \qquad,
 				\label{eq:multipleLevelState}
@@
 
-			\par Being $N=$ the number of levels and $a \in (\alpha, \beta, \gamma, ... , \omega)$ sometimes the equation \ref{eq:multipleLevelState} can be represented as a sum:
+			\par Being $N=$ the number of levels and $a \in (\alpha, \beta, \gamma, \dots , \omega)$ sometimes the equation \ref{eq:multipleLevelState} can be represented as a sum:
 
 			\begin{equation}
-				|\psi\rangle = \sum_{i=1}^{N} a_i . |V_i\rangle = 
+				\ket{\psi} = \sum_{i=1}^{N} a_i . \ket{V_i}= 
 				\begin{bmatrix}
@@
 
-			\par Some authors prefers to use $|0\rangle, |1\rangle, |2\rangle, \dots , |n\rangle / n < N$ instead of $|V_0\rangle,|V_1\rangle, \dots, |V_n\rangle / n < N$.\newline
+			\par Some authors prefers to use $\ket{0}, \ket{1}, \ket{2}, \dots , \ket{n} / n < N$ instead of $\ket{V_0},\ket{V_1}, \dots, \ket{V_n} / n < N$.\newline
 
@@
 				\begin{aligned}
-					|\psi\rangle =
+					\ket{\psi} =
 					\begin{bmatrix}
@@
 
-			\par Briefly, given a ket $|\psi\rangle$ (as shown in equation \ref{eq:ket}), its corresponding bra is denoted by the conjugate transpose of $|\psi\rangle$ (equation \ref{eq:bra}).
+			\par Briefly, given a ket $\ket{\psi}$ (as shown in equation \ref{eq:ket}), its corresponding bra is denoted by the conjugate transpose of $\ket{\psi}$ (equation \ref{eq:bra}).
 
@@
 				\label{eq:ket}
-				|\psi\rangle = 
+				\ket{\psi} = 
 				\begin{bmatrix}
@@
 				\label{eq:bra}
-				\langle\psi| = [u - vo, k - ji, o - li, d - mi]
+				\bra{\psi} = [u - vo, k - ji, o - li, d - mi]
 			\end{equation}
@@
 
-			\par As we already know, a vector of all probability amplitudes, denoted as $|\psi\rangle$, is given by equation \ref{eq:multipleLevelStateKet}.
+			\par As we already know, a vector of all probability amplitudes, denoted as $\bra{\psi}$, is given by equation \ref{eq:multipleLevelStateKet}.
 So, for these "ket" states, there is an associated "bra" notation defined in equation \ref{eq:multipleLevelStateBra} \cite{notacaoDirac}
@@
 			\begin{equation}
-				|\psi\rangle = \sum_{i=1}^{N} a_i . |V_i\rangle = 
+				\ket{\psi} = \sum_{i=1}^{N} a_i . \ket{V_i} = 
 				\begin{bmatrix}
@@
 			\begin{equation}
-				\langle\psi| = \sum_{j=1}^{N} a_j^* . \langle V_j| = [a_1^*, a_2^*, \dots, a_n^*]
+				\bra{\psi} = \sum_{j=1}^{N} a_j^* . \bra{V_j} = [a_1^*, a_2^*, \dots, a_n^*]
 				\label{eq:multipleLevelStateBra}
@@
 				\begin{aligned}
-					\langle\psi|\psi\rangle = [a_1^*, a_2^*, \dots, a_n^*] . \begin{bmatrix}
+					\braket{\psi|\psi} = [a_1^*, a_2^*, \dots, a_n^*] . \begin{bmatrix}
 						a_{1} \\
@@
 			\par Bra notation allows for the calculation of inner products between vectors.
-The inner product of a bra vector $\langle \psi |$ and a ket vector $| \phi \rangle$ is denoted as $\langle \psi | \phi \rangle$.
+The inner product of a bra vector $\bra{\psi}$ and a ket vector $\ket{\phi}$ is denoted as $\braket{\psi|\phi}$.
 The inner product yields a complex number, providing information about the similarity or overlap between the two vectors.\newline
@@
 			\par For instance, given a vector 
-				$|\alpha\rangle= 
+				$\ket{\alpha}= 
 				\begin{bmatrix}
@@
 				 and 
-				 $|\beta\rangle = \begin{bmatrix}
+				 $\ket{\beta} = \begin{bmatrix}
 					0\\
@@
 				\label{eq:braket}
-				\langle \alpha | \beta \rangle = 
+				\braket{\alpha|\beta} = 
 				[1, 0] .
@@
 		\begin{equation}
-			| \psi \rangle = \alpha . | 0 \rangle + \beta . | 1 \rangle = 
+			\ket{\psi} = \alpha . \ket{0} + \beta . \ket{1} = 
 			\alpha . \begin{bmatrix}
@@
 			\beta 
-		\end{bmatrix}$ is the qubit itself and, by convention, $| 1 \rangle = 
+		\end{bmatrix}$ is the qubit itself and, by convention, $\ket{1} = 
 		\begin{bmatrix}
@@
 			1
-		\end{bmatrix} = 1$ and $| 0 \rangle = 
+		\end{bmatrix} = 1$ and $\ket{0} = 
 		\begin{bmatrix}
@@
 
-		\par The Equation \ref{eq:qubit} can be interpreted as follows: The \textbf{non-measured} state of $\psi$ within the vector space defined by vectors $| 1 \rangle$ and $| 0 \rangle$ is a linear combination of them  ($\alpha$ times vector $| 0 \rangle$ plus $\beta$ times vector $| 1 \rangle$).\newline
+		\par The Equation \ref{eq:qubit} can be interpreted as follows: The \textbf{non-measured} state of $\psi$ within the vector space defined by vectors $\ket{1}$ and $\ket{0}$ is a linear combination of them  ($\alpha$ times vector $\ket{0}$ plus $\beta$ times vector $\ket{1}$).\newline
 
@@
 			\par As mentioned earlier, quantum gates are represented as matrices, and their number is infinite because the values of the matrix cells can vary across the real numbers ($\mathbb{R}$).
-Although there is a wide variety of quantum gates and considering $| \psi \rangle = \alpha .
-| 0 \rangle + \beta .
-| 1 \rangle$, here are a few examples.\newline
+Although there is a wide variety of quantum gates and considering $\ket{\psi} = \alpha .
+\ket{0} + \beta .
+\ket{1}$, here are a few examples.\newline
 
@@
 				\label{eq:indentityApplyed}
-				I|\psi\rangle = \begin{bmatrix}
+				I\ket{\psi} = \begin{bmatrix}
 					1& 0 \\
@@
 					\beta
-				\end{bmatrix} = |\psi\rangle
+				\end{bmatrix} = \ket{\psi}
 			\end{equation}
@@
 				\label{eq:flipApplyed}
-				X|\psi\rangle = \begin{bmatrix}
+				X\ket{\psi} = \begin{bmatrix}
 					0& 1 \\
@@
 				\begin{aligned}
-					&H|\psi\rangle = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
+					&H\ket{\psi} = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
 						1& 1 \\
@@
 				\label{eq:rotApplyed}
-				Z|\psi\rangle = \begin{bmatrix}
+				Z\ket{\psi} = \begin{bmatrix}
 					1& 0 \\

### 2023-06-20 documentation/article02/presentation/apresentacao.tex
Presentation ready

@@
 
-\title{Introdução à Computação Quântica}
+\title{Uma breve revisão da Mecânica Quântica para Cientistas da Computação}
 
@@
 
+	\section{Introdução}
+
 	\begin{frame}{Introdução}
+
+		\par Alguns fenômenos, como o \textbf{emaranhamento} e a \textbf{superposição} presentes na mecânica quântica contribuem para uma abordagem mais eficiente em termos de velocidade na execução de alguns algoritmos.
+A computação quântica visa explorar e utilizar esses comportamentos.
+	\end{frame}
+
+	\section{Espaço vetorial}
+
+	\begin{frame}{Espaço vetorial}
+		\par Um \textbf{espaço vetorial} ($V$) sobre os números complexos $\mathbb{C}$ é uma estrutura matemática que permite a representação de seus elementos (vetores) e obedece ao seguinte conjunto de regras:
 		\begin{itemize}
-			\item A mecânica quântica surgiu no século XX, juntamente com os computadores clássicos.
-			\item A computação quântica tem recebido atenção significativa devido ao seu potencial para resolver problemas complexos de forma mais eficiente do que os computadores clássicos.
-			\item A computação quântica utiliza os princípios da mecânica quântica para realizar cálculos usando bits quânticos ou qubits.
+			\item $\alpha \in \mathbb{C}, \alpha . |V_1\rangle \in V$
+			\item $\vec{0} , \vec{0} + |V_1\rangle = |V_1\rangle$ (vetor nulo)
+			\item $|V_1\rangle + |V_2\rangle = |V_2\rangle + |V_1\rangle$ (comutativa)
+			\item $|V_1\rangle \in V, |V_2\rangle \in V \implies |V_1\rangle + |V_2\rangle \in V$
+			\item $|-V_1\rangle; |V_1\rangle + |-V_1\rangle=\vec{0}, |-V_1\rangle=-|V_1\rangle$ 
+			\item $|V_1\rangle + (|V_2\rangle + |V_3\rangle) = (|V_1\rangle + |V_2\rangle) + |V_3\rangle$ (associativa)
+			\item $\alpha,\beta \in \mathbb{C}; \alpha (|V_1\rangle + \beta . |V_2\rangle) = \alpha . |V_1\rangle + \alpha . \beta . |V_2\rangle $ (distributiva)
 		\end{itemize}
@@
 
-	\begin{frame}{Estados Quânticos}
-		\begin{itemize}
-			\item Os estados quânticos são representações matemáticas que fornecem informações sobre as propriedades de um sistema físico, como a posição ou o momento de uma partícula.
-			\item Os estados quânticos são descritos usando vetores de estado, que são frequentemente representados usando a notação de Dirac "ket" (por exemplo, $|\psi\rangle$).
-			\item Ao contrário dos bits clássicos, que podem representar apenas 0 ou 1, os estados quânticos podem existir em uma superposição, onde eles representam simultaneamente vários estados possíveis.
-		\end{itemize}
+	\begin{frame}{Base de um espaço vetorial}
+		\par Sendo $V$ um espaço vetorial, sua base é o conjunto $b$ é tal que $b \subset V$ e os elementos de $b$ são ortonormais, ou seja, ortogonais entre si e unitários. \newline
+		\par Em mecânica quântica esses \textbf{vetores base} são chamados de \textbf{possíveis estados do sistema}.
 	\end{frame}
 
-	\begin{frame}{Espaço Vetorial Complexo}
-		\begin{itemize}
-			\item Os estados quânticos são representados usando espaços vetoriais complexos.
-			\item Um espaço vetorial complexo é uma estrutura matemática que segue regras específicas, incluindo comutatividade e associatividade, permitindo a combinação de vetores.
-			\item Os vetores de base dentro do espaço vetorial representam os estados possíveis do sistema quântico, e as combinações lineares desses vetores de base representam estados quânticos arbitrários.
-		\end{itemize}
+	\section{Notação de Dirac e os estados quânticos}
+
+	\begin{frame}{Notação de Dirac}
+		\par O estado de um sistema quântico é descrito usando um formalismo chamado \textbf{vetores de estado}, que pertencem a um \textbf{espaço vetorial} associado ao sistema.
+Esses vetores são geralmente denotados usando a notação "ket", introduzida pelo físico Paul Dirac, onde um vetor de estado é representado como $|\psi\rangle$, pronunciado como "ket psi".\newline
+		\par A combinação dos símbolos $|$ e $\rangle$ com um valor dentro deles é chamada de "ket". Um exemplo é demonstrado na Equação \ref{eq:ketNotationVector} \cite{notacaoDirac}.
+		\begin{equation}
+			\ket{K} = \vec{K} = \begin{bmatrix}
+				a \\
+				b \\
+				c \\
+				d
+			\end{bmatrix}
+			\label{eq:ketNotationVector}
+		\end{equation}
+		\par Onde $a, b, c, d \in \mathbb{Q}$ e são componentes do \textbf{vetor} (ou matriz coluna) $K$.
 	\end{frame}
 
-	\begin{frame}{Notação de Dirac}
-		\begin{itemize}
-			\item A notação de "BraKet" de Dirac, desenvolvida pelo físico Paul Dirac, é uma notação comum usada na mecânica quântica.
-			\item A notação "ket" ($|\psi\rangle$) representa o estado de uma partícula dentro do espaço vetorial.
-			\item A notação "bra" ($\langle\psi|$) representa o transposto conjugado do vetor ket.
-			\item O produto interno de um bra e um ket fornece a amplitude de probabilidade para um determinado estado.
-		\end{itemize}
+	\begin{frame}[allowframebreaks]{Notação de Dirac para os estados quânticos}
+		\par Para descrever o estado de uma partícula usando \textbf{vetores base}, o estado da partícula pode ser escrito usando a equação \ref{eq:ketNotationGeneric}.
+		\begin{equation}
+			\ket{\psi} = \alpha . \ket{V_1} + \beta . \ket{V_2}
+			\label{eq:ketNotationGeneric}
+		\end{equation}
+		\par Com $\alpha$ e $\beta$ conhecidos como \textbf{amplitudes de probabilidade}.
+
+		\begin{columns}
+			\begin{column}{0.5\textwidth}
+				\begin{figure}[h]
+					\centering
+					\includegraphics{../text/images/diracPlot}
+					\caption{Representação visual da Equação \ref{eq:ketNotationGeneric}. $V_1$ e $V_2$ devem ser ortogonais e normalizados.}
+					\label{fig:diracPlot}
+				\end{figure}
+			\end{column}
+			\begin{column}{0.5\textwidth}
+				\par A figura ao lado mostra espaço vetorial de duas dimensões com bases $V_1$ e $V_2$.
+Mas esse não é o caso da maioria das particulas.
+Um estado quântico bidimensional é apenas um caso particular conveniente para a computação.\newline
+				\par Esse caso particular é chamado de \textbf{sistema de estados em dois níveis}. 
+			\end{column}
+		\end{columns}
+		\par Sendo $N=$ o número de níveis e $a \in (\alpha, \beta, \gamma, \dots , \omega)$ \textbf{sistema de estados multinível} pode ser representada como uma soma:
+		\begin{equation}
+			\ket{\psi} = \sum_{i=1}^{N} a_i . \ket{V_i}= 
+			\begin{bmatrix}
+				a_{1} \\
+				a_{2} \\
+				\vdots \\
+				a_{n}
+			\end{bmatrix}
+			\label{eq:multipleLevelStateSum}
+		\end{equation}
 	\end{frame}
 
-	\begin{frame}{Amplitude de Probabilidade}
-		\begin{itemize}
-			\item As amplitudes de probabilidade são números complexos que descrevem a probabilidade de ocorrência de um estado quântico específico.
-			\item As amplitudes de probabilidade são calculadas usando o produto interno de um bra e um ket.
-			\item As amplitudes de probabilidade são normalizadas para garantir que a soma das probabilidades de todos os estados possíveis seja igual a 1.
-		\end{itemize}
+	\section{Estados Quânticos}
+
+	\begin{frame}{Estados Quânticos}
+		\par Os estados quânticos são representações matemáticas que fornecem informações sobre as propriedades de um sistema físico, como a posição ou o momento de uma partícula.Ao contrário dos bits clássicos, que podem representar apenas 0 ou 1, os estados quânticos podem existir em uma superposição, onde eles representam simultaneamente vários estados possíveis\newline
+
+		\par Os estados quânticos são descritos usando vetores de estado, que são frequentemente representados usando a notação de Dirac "ket" (por exemplo, $|\psi\rangle$).
 	\end{frame}
 
-	\begin{frame}{Operadores Lineares}
-		\begin{itemize}
-			\item Os operadores lineares são ferramentas matemáticas usadas para descrever transformações em espaços vetoriais.
-			\item Na mecânica quântica, os operadores lineares representam observáveis físicos, como posição, momento e energia.
-			\item Os operadores lineares podem ser aplicados a vetores de estado para obter novos vetores de estado resultantes da transformação.
-		\end{itemize}
+	\begin{frame}{Exemplo físico}
+		\par Um bom exemplo do que pode ser um estado quântico são as frequências de onda de um fóton.
+Essa partícula possui componentes elétricos e magnéticos, formando um campo eletromagnético.
+Podemos interpretar, por exemplo, a componente "vertical" como o estado 0 e a "horizontal" como o estado 1, conforme ilustrado na Figura abaixo.
+
+		\begin{columns}
+			\begin{column}{0.5\textwidth}
+				\begin{figure}[H]
+					\centering
+					\includegraphics[width=0.6\linewidth]{../text/images/photonState01}
+					\caption{Estado da partícula é definido pelos valores de amplitude de $\alpha$ e $\beta$ }
+					\label{fig:photonstate01}
+				\end{figure}
+			\end{column}
+			\begin{column}{0.5\textwidth}
+				\par Perceba a semelhança com a representação visual da Equação de Dirac.
+\textbf{Isso não é um coincidência}.
+Neste caso a amplitude das ondas está diretamenta relacionada com o estado dessa partícula.
+			\end{column}
+		\end{columns}
+
 	\end{frame}
 
-	\begin{frame}{Operadores Lineares (cont.)}
-		\begin{itemize}
-			\item Exemplos de operadores lineares comumente usados na mecânica quântica incluem o operador de posição, operador de momento e operador de energia.
-			\item Esses operadores têm propriedades específicas, como hermiticidade e comutatividade, que são importantes na mecânica quântica.
-		\end{itemize}
+	\section{Amplitude de probabilidade}
+
+	\begin{frame}[allowframebreaks]{Amplitude de probabilidade}
+		\par Como já dito anteriomente, sistema de estados em dois níveis, tem como parâmetros definidores os valores de  $\alpha$ e $\beta$. Esses são chamados de amplitude de probabilidade.\newline
+		\par O uso de tais amplitudes se dá no cálculo da probabilidade de um estado colapsar em uma das bases, neste caso $\ket{0}$ ou $\ket{1}$.
+Como tais valores podem estar no domínio dos números complexos $\mathbb{C}$, precisamos introduzir o conceito de conjugado complexo de um número.
+Para um número $z = u + vi$, onde $z \in \mathbb{C}$, seu conjugado é denotado como $z^*$, onde $z^* \in \mathbb{C}$, e é dado por $z^* = u - vi$.
+Em resumo, se $z = u + vi$, então $z^* = u - vi$.
+\newline
+
+		Aplicando o conceito de conjugado complexo, podemos ver que $z \cdot z^* = (u + vi) \cdot (u - vi) = u^2 + v^2 \in \mathbb{R}$.
+Isso resolve o problema, pois o produto $z \cdot z^*$ é um número real!
+\newline
+
+		Finalmente, seja \(P(S_i)\) uma função que calcula a probabilidade do estado \(S_i\) ocorrer.
+Por exemplo, se \(|\psi\rangle = \alpha |S_0\rangle + \beta |S_1\rangle\), então \(P(S_0)\) e \(P(S_1)\) são definidos como mostrado nas equações \ref{eq:state0} e \ref{eq:state1} \cite{binney2013physics}.
+
+		\begin{equation}
+			\label{eq:state0}
+			P(S_0) = \alpha . \alpha^* 
+		\end{equation}
+
+		\begin{equation}
+			\label{eq:state1}
+			P(S_1) = \beta . \beta^* 
+		\end{equation}
+
 	\end{frame}
 
-	\begin{frame}{Portas Quânticas}
-		\begin{itemize}
-			\item Portas quânticas são análogas aos circuitos lógicos em computadores clássicos.
-			\item Elas são operadores lineares que agem em qubits individuais ou em múltiplos qubits.
-			\item As portas quânticas permitem a transformação dos estados quânticos, possibilitando várias operações computacionais.
-			\item Algumas portas quânticas conhecidas incluem a porta Hadamard, porta Pauli-X e a porta CNOT.
-		\end{itemize}
+	\section{Princípio de Heisenberg}
+
+	\begin{frame}[allowframebreaks]{Não se pode ter tudo}
+		\par O princípio de Heisenberg se baseia no fato que a medição simultânea de certas grandezas nem sempre é posível: Se sabemos muito sobre a velocidade de uma partícula, pouco se sabe sobre sua posição; se é certa a presença de uma frequência em um sinal é difil saber quando a mesma ocorre, etc.
+		\par O último exemplo supracitado é ilustrado na figuras abaixo:
+
+		\begin{columns}
+			\begin{column}{0.5\textwidth}
+				\begin{figure}[h]
+					\centering
+					\includegraphics[width=1\linewidth]{../text/images/noisySignal}
+					\caption[Multi-frequency signal]{Multi-frequency signal. Source: \cite{olama2011design}}
+					\label{fig:noisysignal}
+				\end{figure}
+			\end{column}
+			\begin{column}{0.5\textwidth}
+					\begin{figure}[h]
+						\centering
+						\includegraphics[width=0.1\linewidth]{../text/images/windowedNoisySignal}
+						\caption[Windowed signal]{Small windowed signal. Adapted from: \cite{olama2011design}}
+						\label{fig:windowednoisysignal}
+					\end{figure}
+			\end{column}
+		\end{columns}
+		\par Perceba que, quando há acesso ao sinal por inteiro é fácil via uma transfomada de Fourrier, por exemplo, saber quais a frequências envolvidas no sinal, o mesmo não acontece qundo o foco é em alguma região de alta frequência.
 	\end{frame}
 
-	\begin{frame}{Portas Quânticas (cont.)}
+	\section{Qubit ou bit quântico}
+
+	\begin{frame}[allowframebreaks]{Qubit ou bit quântico}
+		\par Qubits existem em uma superposição dos estados 1 e 0 até serem medidos, momento em que seu estado colapsa para 1 ou 0 com probabilidades dadas por $|\alpha|^ 2$ e $|\beta|^2$ \cite{da2018introduccao}.
+
+		\begin{equation}
+			\ket{\psi} = \alpha . \ket{0} + \beta . \ket{1} = 
+			\alpha . \begin{bmatrix}
+				1 \\
+				0
+			\end{bmatrix} + \beta .	\begin{bmatrix}
+				0 \\
+				1
+			\end{bmatrix} = 
+			\begin{bmatrix}
+				\alpha \\
+				\beta
+			\end{bmatrix}
+			\label{eq:qubit}
+		\end{equation}
+
+		\begin{columns}
+			\begin{column}{.5\linewidth}
+				\begin{figure}
+					\centering
+					\includegraphics[width=0.5\linewidth]{../text/images/photonState02}
+					\caption{Ao ser lido este qubit tem maior probabilidade de colapsar em 0.}
+					\label{fig:photonstate02}
+				\end{figure}
+			\end{column}
+			\begin{column}{.5\linewidth}
+				\begin{figure}
+					\centering
+					\includegraphics[width=0.5\linewidth]{../text/images/photonState03}
+					\caption{Ao ser lido este qubit tem maior probabilidade de colapsar em 1.}
+					\label{fig:photonstate03}
+				\end{figure}
+			\end{column}
+		\end{columns}
+
+		\par Qubits têm a capacidade de armazenar e processar informações de maneira diferente dos bits clássicos devido à superposição.
+A sua capacidade, em termos de armazenamento de informação, é significativamente maior em comparação com os bits clássicos.
+No entanto, há um problema: quando os qubits são medidos, eles se transformam em bits clássicos.
+Portanto, atualmente, não é possível aproveitar ao máximo as propriedades quânticas para fins de armazenamento de dados.
+Além disso, os qubits podem armazenar dados não binários, como	 $\begin{bmatrix} \dfrac{1}{\sqrt{2}} \\\\ \dfrac{1}{\sqrt{2}} \end{bmatrix}$ or $\begin{bmatrix} \dfrac{1}{2} \\\\ \dfrac{\sqrt{3}}{2} \end{bmatrix}$.
+	\end{frame}
+
+	\begin{frame}[allowframebreaks]{Portas Quânticas}
 		\begin{itemize}
-			\item Porta Hadamard: Ela cria uma superposição de estados em um qubit. É representada pela seguinte matriz:
-			\[
-			H = \frac{1}{\sqrt{2}}\begin{bmatrix}
+			\item Porta identidade: Mantém o qubit no mesmo estado.
+			$I : \begin{bmatrix}
+					1& 0 \\
+					0& 1
+				\end{bmatrix}$
+
+			\item Porta Hadamard: Ela cria uma superposição de estados em um qubit. É representada pela seguinte matriz \cite{qcfcs}:
+			$H = \frac{1}{\sqrt{2}}\begin{bmatrix}
 				1 & 1 \\
 				1 & -1 \\
-			\end{bmatrix}
-			\]
-			\item Porta Pauli-X: Conhecida como porta NOT quântica, ela inverte o estado de um qubit. É representada pela seguinte matriz:
-			\[
-			X = \begin{bmatrix}
+			\end{bmatrix}$
+			\item Porta X, ela inverte o estado de um qubit. É representada por:
+			$X = \begin{bmatrix}
 				0 & 1 \\
 				1 & 0 \\
-			\end{bmatrix}
-			\]
+			\end{bmatrix}$
 			\item Porta CNOT: A porta Controlled NOT opera em dois qubits, onde um atua como controle e o outro como alvo. É representada pela seguinte matriz:
-			\[
-			CNOT = \begin{bmatrix}
+			$CNOT = \begin{bmatrix}
 				1 & 0 & 0 & 0 \\
@@
 				0 & 0 & 1 & 0 \\
+			\end{bmatrix}$
+		\end{itemize}
+
+		\par Este portão opera em \textbf{dois qubits}, onde um é o \textbf{qubit de controle} e o outro é o \textbf{target qubit}.
+Para representar o estado dos qubits combinados, é utilizada uma variação da notação ket, onde dois vetores são representados juntos.
+Isso é necessário porque os qubits não podem mais ser escritos separadamente devido ao emaranhamento que esse portão pode causar.
+As variações do ket são mostradas nas Equações \ref{eq:ketForTwoVectors00}, \ref{eq:ketForTwoVectors01} e \ref{eq:ketForTwoVectors02}.
+
+		\begin{equation}
+			\label{eq:ketForTwoVectors00}
+			\ket{1}, \ket{0} \implies
+			\ket{10} = \begin{bmatrix}
+				0 \\
+				1 \\
+				1 \\
+				0
 			\end{bmatrix}				
-			\]
-		\end{itemize}
-	\end{frame}
+		\end{equation}
 
-	\begin{frame}{Conclusão}
-		\begin{itemize}
-			\item A computação quântica é um campo promissor que utiliza os princípios da mecânica quântica para realizar cálculos mais eficientes.
-			\item Os estados quânticos, espaços vetoriais complexos, operadores lineares e portas quânticas são conceitos fundamentais na computação quântica.
-			\item As portas quânticas permitem a realização de operações computacionais, e exemplos conhecidos incluem a porta Hadamard, porta Pauli-X e porta CNOT.
-			\item A compreensão desses conceitos é essencial para explorar o potencial da computação quântica e suas aplicações futuras.
-		\end{itemize}
+		\begin{equation}
+			\label{eq:ketForTwoVectors01}
+			\ket{0}, \ket{0} \implies
+			\ket{00} = \begin{bmatrix}
+				1 \\
+				0 \\
+				1 \\
+				0
+			\end{bmatrix}				
+		\end{equation}
+
+		\begin{equation}
+			\label{eq:ketForTwoVectors02}
+			\ket{1}, \ket{1} \implies
+			\ket{11} = \begin{bmatrix}
+				0 \\
+				1 \\
+				0 \\
+				1
+			\end{bmatrix}				
+		\end{equation}
+
+		\par O portão CNOT opera em dois qubits, onde o primeiro qubit serve como controle e o segundo qubit como alvo.
+Conforme mostrado na Equação \ref{eq:cnotApplyed}, quando o qubit de controle está no estado $\ket{1}$, o qubit alvo sofre uma inversão.
+No entanto, quando o qubit de controle está no estado $\ket{0}$, o qubit de destino permanece inalterado \cite{nielsen2002quantum}.
+
+		\begin{equation}
+			\label{eq:cnotApplyed}
+			\begin{aligned}
+				&CNOT\ket{00} = \ket{00}, \\
+				&CNOT\ket{01} = \ket{01}, \\
+				&CNOT\ket{10} = \ket{11}, \\
+				&CNOT\ket{11} = \ket{10}
+			\end{aligned}
+		\end{equation}
 	\end{frame}
@@
 		\frametitle{Referências}
-		\bibliography{bibliography.bib}
+		\bibliography{../text/references.bib}
 	\end{frame}

### 2023-06-20 code2/IntegralEvaluator/main.cpp

@@ -1,0 +1,81 @@
+#include <cmath>
+#include <iomanip>
+#include <iostream>
+#include <mpi.h>
+
+/*
+ * Inlining functions so the call overhead
+ * is avoided
+ */
+inline long double f(long double x)
+{
+    return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
+}
+
+/*
+ * Inlining the functions so the call overhead
+ * is avoided. Note that **inline** is not a
+ * command; the compiler may **ignore** the inline
+ * request.
+ */
+inline long double integrate(long double (*f)(long double),
+                             double lowerLimit,
+                             double upperLimit,
+                             int partitions)
+{
+    double h = (upperLimit - lowerLimit) / partitions;
+    long double sum = 0;
+
+    for (int k = 1; k < partitions - 1; k++) {
+        sum += f(lowerLimit + k * h);
+    }
+
+    sum *= 2;
+    sum += (f(lowerLimit) + f(upperLimit));
+    return (h / 2) * sum;
+}
+
+inline void experiment(
+    double lowerLimit, double upperLimit, double interval, int myRank, int numProcs)
+{
+    long double result = 0;
+    int partitions = static_cast<int>(upperLimit / interval);
+    int localPartitions = partitions / numProcs;
+    int localStart = myRank * localPartitions;
+    int localEnd = localStart + localPartitions;
+
+    long double localSum = integrate(f,
+                                     lowerLimit + localStart * interval,
+                                     lowerLimit + localEnd * interval,
+                                     localPartitions);
+
+    // Gather results from all processes
+    MPI_Reduce(&localSum, &result, 1, MPI_LONG_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD);
+
+    if (myRank == 0) {
+        std::cout << std::fixed << std::setprecision(6) << interval << " \t " << partitions
+                  << " \t " << result << std::endl;
+    }
+}
+
+int main(int argc, char **argv)
+{
+    double lowerLimit = 0, upperLimit = 100;
+    int numProcs, myRank;
+
+    MPI_Init(&argc, &argv);
+    MPI_Comm_size(MPI_COMM_WORLD, &numProcs);
+    MPI_Comm_rank(MPI_COMM_WORLD, &myRank);
+
+    if (myRank == 0) {
+        std::cout << "Interval \t Partitions \t Result" << std::endl;
+    }
+
[…]

### 2023-07-10 documentation/article02/presentation/apresentacao.tex
Open MPI version implemented

@@
 Além disso, os qubits podem armazenar dados não binários, como	 $\begin{bmatrix} \dfrac{1}{\sqrt{2}} \\\\ \dfrac{1}{\sqrt{2}} \end{bmatrix}$ or $\begin{bmatrix} \dfrac{1}{2} \\\\ \dfrac{\sqrt{3}}{2} \end{bmatrix}$.
+
+		\framebreak
+
+		\par A lista abaixo mostra a capacidade dos qubits em compração com os bits clássicos:
+
+		\begin{itemize}
+			\label{lst:qubits}
+			\item 1 qubit = 2 bits
+			\item 2 qubits = 4 bits
+			\item 3 qubits = 8 bits (1 byte)
+			\item 4 qubits = 16 bits
+			\item 13 qubits = 8,192 bits (1 kilobyte)
+			\item 23 qubits = 8,388,608 bits (1 megabyte)
+			\item 33 qubits = 8,589,934,592 bits (1 gigabyte)
+			\item 43 qubits = 8,796,093,022,208 bits (1 terabyte)
+			\item n qubits = $2^n$ bits
+		\end{itemize}
 	\end{frame}
@@
 
-		\par Este portão opera em \textbf{dois qubits}, onde um é o \textbf{qubit de controle} e o outro é o \textbf{target qubit}.
+		\par Esta porta opera em \textbf{dois qubits}, onde um é o \textbf{qubit de controle} e o outro é o \textbf{target qubit}.
 Para representar o estado dos qubits combinados, é utilizada uma variação da notação ket, onde dois vetores são representados juntos.
-Isso é necessário porque os qubits não podem mais ser escritos separadamente devido ao emaranhamento que esse portão pode causar.
+Isso é necessário porque os qubits não podem mais ser escritos separadamente devido ao emaranhamento que essa porta pode causar.
 As variações do ket são mostradas nas Equações \ref{eq:ketForTwoVectors00}, \ref{eq:ketForTwoVectors01} e \ref{eq:ketForTwoVectors02}.
@@
 
-		\par O portão CNOT opera em dois qubits, onde o primeiro qubit serve como controle e o segundo qubit como alvo.
+		\par A porta CNOT opera em dois qubits, onde o primeiro qubit serve como controle e o segundo qubit como alvo.
 Conforme mostrado na Equação \ref{eq:cnotApplyed}, quando o qubit de controle está no estado $\ket{1}$, o qubit alvo sofre uma inversão.

### 2023-07-10 code2/IntegralEvaluator/main.cpp

@@ -4,7 +4,7 @@
 #include <mpi.h>
 
 /*
- * Inlining functions so the call overhead
+ * Inlinning functions so the call overhead
  * is avoided
  */
 inline long double f(long double x)
@@ -13,9 +13,9 @@ inline long double f(long double x)
 }
 
 /*
- * Inlining the functions so the call overhead
- * is avoided. Note that **inline** is not a
- * command; the compiler may **ignore** the inline
+ * Inlinning the functions so the call overhead
+ * is avoided. Note that **inline** it is not a
+ * command, the compiler may **ignore** the inline
  * request.
  */
 inline long double integrate(long double (*f)(long double),
@@ -35,45 +35,54 @@ inline long double integrate(long double (*f)(long double),
     return (h / 2) * sum;
 }
 
-inline void experiment(
-    double lowerLimit, double upperLimit, double interval, int myRank, int numProcs)
+inline void experiment(double lowerLimit, double upperLimit, double interval)
 {
     long double result = 0;
-    int partitions = static_cast<int>(upperLimit / interval);
-    int localPartitions = partitions / numProcs;
-    int localStart = myRank * localPartitions;
-    int localEnd = localStart + localPartitions;
+    int partitions = upperLimit / interval;
 
-    long double localSum = integrate(f,
-                                     lowerLimit + localStart * interval,
-                                     lowerLimit + localEnd * interval,
-                                     localPartitions);
+    double start = MPI_Wtime();
+    result = integrate(f, lowerLimit, upperLimit, partitions);
+    double end = MPI_Wtime();
 
-    // Gather results from all processes
-    MPI_Reduce(&localSum, &result, 1, MPI_LONG_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD);
+    // Calculating total time taken by the program.
+    double timeTaken = end - start;
 
-    if (myRank == 0) {
+    int rank, size;
+    MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+    MPI_Comm_size(MPI_COMM_WORLD, &size);
+
+    long double *allResults = nullptr;
+
+    if (rank == 0) {
+        allResults = new long double[size];
+    }
+
[…]

### 2023-07-11 code3/main.cpp
Forgotten images added

@@ -1,0 +1,97 @@
+#include <CL/cl.hpp>
+#include <cmath>
+#include <iomanip>
+#include <iostream>
+#include <vector>
+
+/*
+ * Inlinning functions so the call overhead
+ * is avoided
+ */
+inline long double f(long double x)
+{
+    return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
+}
+
+inline long double integrate(std::vector<long double> &f_values,
+                             double lowerLimit,
+                             double upperLimit,
+                             int partitions)
+{
+    double h = (upperLimit - lowerLimit) / partitions;
+    long double sum = 0;
+
+    // Create an OpenCL context and queue
+    cl::Context context(CL_DEVICE_TYPE_DEFAULT);
+    std::vector<cl::Device> devices = context.getInfo<CL_CONTEXT_DEVICES>();
+    cl::CommandQueue queue(context, devices[0]);
+
+    // Load the OpenCL kernel source code
+    std::string kernelSource = "__kernel void integrate(__global float* f_values, float "
+                               "lowerLimit, float h, int partitions) {\n"
+                               "    int i = get_global_id(0);\n"
+                               "    float x = lowerLimit + i * h;\n"
+                               "    f_values[i] = sqrt(pow(100.0f, 2) - pow(x, 2));\n"
+                               "}\n";
+
+    // Create an OpenCL program and build it
+    cl::Program::Sources sources(1, std::make_pair(kernelSource.c_str(), kernelSource.length() + 1));
+    cl::Program program(context, sources);
+    program.build(devices);
+
+    // Create an OpenCL kernel
+    cl::Kernel kernel(program, "integrate");
+
+    // Create OpenCL buffers for f_values and sum
+    cl::Buffer fBuffer(context, CL_MEM_WRITE_ONLY, sizeof(long double) * partitions);
+
+    // Set kernel arguments
+    kernel.setArg(0, fBuffer);
+    kernel.setArg(1, static_cast<float>(lowerLimit));
+    kernel.setArg(2, static_cast<float>(h));
+    kernel.setArg(3, partitions);
+
+    // Enqueue the kernel for execution
+    queue.enqueueNDRangeKernel(kernel, cl::NullRange, cl::NDRange(partitions));
+
+    // Read the results back to the host
+    queue.enqueueReadBuffer(fBuffer, CL_TRUE, 0, sizeof(long double) * partitions, f_values.data());
+
[…]

### 2023-07-11 documentation/article03/article.tex
Open MPI implementation

--- 2023-06-09 documentation/article01/article.tex
@@
 \begin{document}
-	\title{A tutorial of Solving the defined integral problem with OpenMp}
+	\title{A tutorial of Solving the defined integral problem with OpenMp and Open MPI}
 	\author{André Furlan - UNESP - Universidade Estadual Paulista "Júlio de Mesquita Filho"}
@@
 	\begin{abstract}
-		The Trapezoid rule algorithm used in this tutorial is \textbf{not} the most fast method to calculate an interval defined integral but is a good example of how to make programs runs quicker by using OpenMp for parallelism.
-Hopefully it will help the understanding of how this library works and how to use it.
+		The Trapezoid rule algorithm used in this tutorial is \textbf{not} the most fast method to calculate an interval defined integral but is a good example of how to make programs runs quicker by using OpenMp or Open MPI for parallelism.
+Hopefully it will help the understanding of how these libraries works and how to use them.
 All sources can be accessed at \href{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}{https://github.com/ensismoebius/ComputacaoDeAltoDesempenho}
@@
 	\begin{IEEEkeywords}
-		OpenMp, Paralelism, C++, C, Tutorial
+		OpenMp, Open MPI, Paralelism, C++, C, Tutorial
 	\end{IEEEkeywords}
@@
 Most of the time this means use all processors an its cores at maximum, avoiding idle intervals.
-	\par By using the OpenMp \cite{openmp08} libraries it is possible, using \textbf{\#pragma} compiler directives, turn a sequential program into a parallel one.
-The problem in focus here is the defined interval integral calculation.
+	\par By using the OpenMp \cite{openmp08} library it is possible by using \textbf{\#pragma} compiler directives, turn a sequential program into a parallel one.
+
+	\par When using the Open MPI library, \cite{mpi40} it is necessary to use the "MPI\_" prefixed commands.
+The "MPI\_" prefix is a convention used in Open MPI implementations, but it is not required by the MPI standard.
+
+	\par The problem in focus here is the defined interval integral calculation.
 To solve it the program must calculate the area bellow the plot by summing up the area of several trapezoids \cite{hildebrand1987introduction}.
@@
 And it is here that the parallelism enters: By dividing the work of area calculation in lots of threads its possible to diminish the effective time of execution.
+
 	\section{The trapezoid rule}
@@
 
-		\par Then, for the impatient ones, here is the complete code that will be discussed:
+		\subsection{OpenMp implementation}	
 			\lstinputlisting[language=C++]{code/integral.cpp}
+
+		\subsection{Open MPI implementation}	
+			\lstinputlisting[language=C++]{code/integral2.cpp}
 

### 2023-07-11 documentation/article03/code/integral2.cpp.tex
Open MPI implementation

#include <cmath>
#include <cstddef>
#include <iomanip>
#include <iostream>
#include <mpi.h>
#include <ostream>

inline long double f(long double x)
{
	return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
}

inline long double integrate(long double (*f)(long double),
double lowerLimit,
double upperLimit,
int partitions)
{
	double h = (upperLimit - lowerLimit) / partitions;
	long double sum = 0;

	int slotId, slotsAvailable;
	MPI_Comm_rank(MPI_COMM_WORLD, &slotId);
	MPI_Comm_size(MPI_COMM_WORLD, &slotsAvailable);

	int localIterations = (partitions - 2) / slotsAvailable;
	int remainingIterations = (partitions - 2) % slotsAvailable;

	int start = slotId * localIterations + std::min(slotId, remainingIterations) + 1;
	int end = start + localIterations + (slotId < remainingIterations ? 1 : 0);

	long double localSum = 0;

	for (int k = start; k < end; k++) {
		localSum += f(lowerLimit + k * h);
	}

	MPI_Reduce(&localSum, &sum, 1, MPI_LONG_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD);

	if (slotId == 0) {
		sum *= 2;
		sum += (f(lowerLimit) + f(upperLimit));
		sum *= (h / 2);
	}

	return sum;
}

inline void experiment(double lowerLimit, double upperLimit, double interval)
{
	long double result = 0;
	int partitions = upperLimit / interval;

	double start_time = MPI_Wtime();

	result = integrate(f, lowerLimit, upperLimit, partitions);

	double end_time = MPI_Wtime();
	double elapsed_time = end_time - start_time;

	int rank;
	MPI_Comm_rank(MPI_COMM_WORLD, &rank);

	if (rank == 0) {
		std::cout << std::fixed << std::setprecision(6) << interval << " \t " << partitions
		<< " \t " << elapsed_time << " \t " << result << std::endl;
	}
}

int main(int argc, char *argv[])
{
	MPI_Init(NULL, NULL);

	double lowerLimit = 0, upperLimit = 100;

	int slotId;
	MPI_Comm_rank(MPI_COMM_WORLD, &slotId);

	if (slotId == 0) {
		std::cout << "Interval \t Partitions \t Time taken \t Result" << std::endl;
	}

	experiment(lowerLimit, upperLimit, 0.000001);
	experiment(lowerLimit, upperLimit, 0.00001);
	experiment(lowerLimit, upperLimit, 0.0001);

	MPI_Finalize();
	return 0;
}

### 2023-07-11 code2/main.cpp

@@ -1,23 +1,15 @@
 #include <cmath>
+#include <cstddef>
 #include <iomanip>
 #include <iostream>
 #include <mpi.h>
+#include <ostream>
 
-/*
- * Inlinning functions so the call overhead
- * is avoided
- */
 inline long double f(long double x)
 {
     return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
 }
 
-/*
- * Inlinning the functions so the call overhead
- * is avoided. Note that **inline** it is not a
- * command, the compiler may **ignore** the inline
- * request.
- */
 inline long double integrate(long double (*f)(long double),
                              double lowerLimit,
                              double upperLimit,
@@ -26,13 +18,31 @@ inline long double integrate(long double (*f)(long double),
     double h = (upperLimit - lowerLimit) / partitions;
     long double sum = 0;
 
-    for (int k = 1; k < partitions - 1; k++) {
-        sum += f(lowerLimit + k * h);
+    int slotId, slotsAvailable;
+    MPI_Comm_rank(MPI_COMM_WORLD, &slotId);
+    MPI_Comm_size(MPI_COMM_WORLD, &slotsAvailable);
+
+    int localIterations = (partitions - 2) / slotsAvailable;
+    int remainingIterations = (partitions - 2) % slotsAvailable;
+
+    int start = slotId * localIterations + std::min(slotId, remainingIterations) + 1;
+    int end = start + localIterations + (slotId < remainingIterations ? 1 : 0);
+
+    long double localSum = 0;
+
+    for (int k = start; k < end; k++) {
+        localSum += f(lowerLimit + k * h);
     }
 
-    sum *= 2;
-    sum += (f(lowerLimit) + f(upperLimit));
-    return (h / 2) * sum;
+    MPI_Reduce(&localSum, &sum, 1, MPI_LONG_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD);
+
+    if (slotId == 0) {
+        sum *= 2;
+        sum += (f(lowerLimit) + f(upperLimit));
+        sum *= (h / 2);
+    }
+
+    return sum;
 }
 
 inline void experiment(double lowerLimit, double upperLimit, double interval)
@@ -40,43 +50,32 @@ inline void experiment(double lowerLimit, double upperLimit, double interval)
     long double result = 0;
     int partitions = upperLimit / interval;
 
[…]

### 2023-07-11 code3/main.cpp

@@ -1,97 +1,0 @@
-#include <CL/cl.hpp>
-#include <cmath>
-#include <iomanip>
-#include <iostream>
-#include <vector>
-
-/*
- * Inlinning functions so the call overhead
- * is avoided
- */
-inline long double f(long double x)
-{
-    return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
-}
-
-inline long double integrate(std::vector<long double> &f_values,
-                             double lowerLimit,
-                             double upperLimit,
-                             int partitions)
-{
-    double h = (upperLimit - lowerLimit) / partitions;
-    long double sum = 0;
-
-    // Create an OpenCL context and queue
-    cl::Context context(CL_DEVICE_TYPE_DEFAULT);
-    std::vector<cl::Device> devices = context.getInfo<CL_CONTEXT_DEVICES>();
-    cl::CommandQueue queue(context, devices[0]);
-
-    // Load the OpenCL kernel source code
-    std::string kernelSource = "__kernel void integrate(__global float* f_values, float "
-                               "lowerLimit, float h, int partitions) {\n"
-                               "    int i = get_global_id(0);\n"
-                               "    float x = lowerLimit + i * h;\n"
-                               "    f_values[i] = sqrt(pow(100.0f, 2) - pow(x, 2));\n"
-                               "}\n";
-
-    // Create an OpenCL program and build it
-    cl::Program::Sources sources(1, std::make_pair(kernelSource.c_str(), kernelSource.length() + 1));
-    cl::Program program(context, sources);
-    program.build(devices);
-
-    // Create an OpenCL kernel
-    cl::Kernel kernel(program, "integrate");
-
-    // Create OpenCL buffers for f_values and sum
-    cl::Buffer fBuffer(context, CL_MEM_WRITE_ONLY, sizeof(long double) * partitions);
-
-    // Set kernel arguments
-    kernel.setArg(0, fBuffer);
-    kernel.setArg(1, static_cast<float>(lowerLimit));
-    kernel.setArg(2, static_cast<float>(h));
-    kernel.setArg(3, partitions);
-
-    // Enqueue the kernel for execution
-    queue.enqueueNDRangeKernel(kernel, cl::NullRange, cl::NDRange(partitions));
-
-    // Read the results back to the host
-    queue.enqueueReadBuffer(fBuffer, CL_TRUE, 0, sizeof(long double) * partitions, f_values.data());
-
[…]

### 2023-07-11 documentation/article03/code/integral.cpp

@@ -1,0 +1,58 @@
+#include <cmath>
+#include <iostream>
+#include <ostream>
+
+#ifdef _OPENMP
+	#include <omp.h>
+#endif
+
+inline long double f(long double x)
+{
+    return std::sqrt(std::pow(100, 2) - std::pow(x, 2));
+}
+
+inline long double integrate(long double (*f)(long double),
+                             double lowerLimit,
+                             double upperLimit,
+                             int partitions)
+{
+    double h = (upperLimit - lowerLimit) / partitions;
+    long double sum = 0;
+
+#pragma default(none) shared(soma)
+    {
+#pragma omp parallel
+        {
+#pragma omp for reduction(+ : sum)
+            for (int k = 1; k < partitions - 1; k++) {
+                sum += f(lowerLimit + k * h);
+            }
+        }
+#pragma omp barrier
+
+        sum *= 2;
+        sum += (f(lowerLimit) + f(upperLimit));
+        return (h / 2) * sum;
+    }
+}
+
+int main()
+{
+    double lowerLimit = 0, upperLimit = 100;
+
+    int partitions = 0;
+
+    partitions = upperLimit / 0.000001;
+    std::cout << "Interval: " << 0.000001 << " Partitions: " << partitions
+              << " - Result: " << integrate(f, lowerLimit, upperLimit, partitions) << std::endl;
+
+    partitions = upperLimit / 0.00001;
+    std::cout << "Interval: " << 0.00001 << " Partitions: " << partitions
+              << " - Result: " << integrate(f, lowerLimit, upperLimit, partitions) << std::endl;
+
+    partitions = upperLimit / 0.0001;
+    std::cout << "Interval: " << 0.0001 << " Partitions: " << partitions
+              << " - Result: " << integrate(f, lowerLimit, upperLimit, partitions) << std::endl;
+
+    return 0;
+}

### 2023-07-11 code2/runmpi.sh
Bash file to run an openmpi executable

@@ -1,0 +1,2 @@
+# Change ./yourExecutable to the path of your open MPI executable
+mpirun --hostfile hosts.txt -np 10 ./yourExecutable

### 2023-07-11 code2/runmpi.sh
A minor fix

@@ -1,2 +1,2 @@
-# Change ./yourExecutable to the path of your open MPI executable
-mpirun --hostfile hosts.txt -np 10 ./yourExecutable
+# Change "build-IntegralEvaluator-Desktop-Release/parallel_integration" to the path of your open MPI executable
+mpirun --hostfile hosts.txt -np 10 ./build-IntegralEvaluator-Desktop-Release/parallel_integration

### 2023-07-11 documentation/article03/article.tex
Open MPI part is done!

@@
 	\section{Discussion}
-		\par This section focus only in the OpenMp directives, its assumed that the reader knows a minimum of C++.\newline
+		\par This section focus only in the OpenMp and Open MPI directives, its assumed that the reader knows a minimum of C++.\newline
+
+		\subsection{OpenMp discussion}
 
@@
 
+		\subsection{Open MPI}
+		\par On line 4 there is a \textbf{required} include for the Open MPI library.
+		\begin{lstlisting}[language=C++]
+#include <mpi.h>
+		\end{lstlisting}
+
+		\par From line 6 to 9, the function to be integrated is defined.
+Note the \textit{inline} statement: Inlinning functions \textbf{may} avoid the intrinsic overhead of calling a function but, since this it is not a command, the compiler may ignore this request.
+
+		\begin{lstlisting}[language=C++]
+inline long double f(long double x)
+{
+	return std::sqrt(std::pow(100,2) - std::pow(x,2));
+}
+		\end{lstlisting}
+
+		\par On the nice 69 line there is a command that initializes the MPI environment, including communication channels and resources necessary for running parallel Open MPI programs.
+The "MPI\_Init" command should be called exactly once per MPI process and has the same arguments as the C++ main function i.e.
+the program arguments count and the arguments.
+
+		\begin{lstlisting}[language=C++]
+MPI_Init(&argc, &argv);
+		\end{lstlisting}
+
+		\par At line 73 and 74 there is a slot identificantion retrieval.
+A slot is a resource to make computation (usually an CPU core) locally or remotely available.
+This slot information will be used in the next lines.
+
+	\begin{lstlisting}[language=C++]
+int slotId;
+MPI_Comm_rank(MPI_COMM_WORLD, &slotId);
+	\end{lstlisting}
+
+	\par On line 76 to 78 the intended message is only shown if the process is from slot 0 i.e. the main process.
+	\begin{lstlisting}[language=C++]
+ if (slotId == 0) {
+	std::cout << "Interval \t Partitions \t Time taken \t Result" << std::endl;
+}
+	\end{lstlisting}
+
+	\par The line 51 begins the time measurement.
+
+	\begin{lstlisting}[language=C++]
+double start_time = MPI_Wtime();
+	\end{lstlisting}
+
+	\par Line 55 and 56 measure time again and calculates the elapsed time respectively.
+
+	\begin{lstlisting}[language=C++]
+double end_time = MPI_Wtime();
+double elapsed_time = end_time - start_time;
+	\end{lstlisting}
+
+	\par As already explained from lines 58 to 64 it guarantees that the elapsed times outputs only runs on main slot.
+
+	\begin{lstlisting}[language=C++]
+int rank;
+MPI_Comm_rank(MPI_COMM_WORLD, &rank);
+
+if (rank == 0) {
+	std::cout << std::fixed << std::setprecision(6) << interval << " \t " << partitions
+	<< " \t " << elapsed_time << " \t " << result << std::endl;
+}
+	\end{lstlisting}
+
+	\par Line 19 creates the slots variables (\textit{slotId} as the current slot/process identification and \textit{slotsAvailable} as the number of available slots).
+Lines 20 and 21 retrieves these values.
+
+	\begin{lstlisting}[language=C++]
+int slotId, slotsAvailable;
+MPI_Comm_rank(MPI_COMM_WORLD, &slotId);
+MPI_Comm_size(MPI_COMM_WORLD, &slotsAvailable);
+	\end{lstlisting}
+
+	\par The line 35 aggregates all sums from \textit{localSum} variables across the slots in the \textit{sum} variable.
+
+	\begin{lstlisting}[language=C++]
+MPI_Reduce(&localSum, &sum, 1, MPI_LONG_DOUBLE, MPI_SUM, 0, MPI_COMM_WORLD);
+	\end{lstlisting}
+
+	\par Finally at 84 line all resources allocated are freed.
+	\begin{lstlisting}[language=C++]
+MPI_Finalize();
+	\end{lstlisting}
+
+	\par Note that most of the processing is done in lines 31 to 33, and the problem \textbf{must} be partitioned manually in Open MPI, which is different from OpenMP.
+
 	\section{Performance tests}
 		\par The tests were done using an \textit{AMD Ryzen 5 5000 series} processor running an \textit{Gnu/Linux operational system} with \textit{30GB of RAM}.\newline
+
+		\subsection{OpenMp performance tests}
 
@@
 
+	\section{Performance tests}
+\par The tests were done using an \textit{AMD Ryzen 5 5000 series} processor running an \textit{Gnu/Linux operational system} with \textit{30GB of RAM}.\newline
+
+\subsection{Open MPI performance tests}
+
+\par In order to change the number of parallel processing units in an Open MPI process, first you need to create a text file that contains the "hosts" and their respective available slots that will be used by Open MPI.
+The hosts can be specified using their DNS name or IP address.
+In this example a file named "hosts.txt" was created.
+
+\begin{lstlisting}
+localhost slots=8
+\end{lstlisting}
+
+\par But it could be like the one bellow:
+
+\begin{lstlisting}
+localhost slots=4
+192.168.0.56 slots=10
+192.168.0.101 slots=7
+\end{lstlisting}
+
+\par When the aforementioned file is created, you can start the processing using the command bellow.
+Note the use of the "-np" parameter, which represents the number of slots to be used and can be interpreted as the number of threads in Open MPI.
+If the specified "-np" value is greater than the number of available slots, an error will be raised.
+
+\begin{lstlisting}
+mpirun --hostfile hosts.txt -np 1 ./executable
+\end{lstlisting}
+
+\par The tables bellow shows the results according to the number of slots.
+
+\begin{table}[h]
+	\caption{mpirun --hostfile hosts.txt -np 1}
+	\begin{center}
+		\centering
+		\begin{tabular}{|l|l|l|l|}
+			\hline
+			Interval & Partitions & Time taken & Result \\
+			\hline
+			0.000001 & 100000000 & 0.470636 & 7853.981634 \\
+			0.000010 & 10000000 & 0.044703 & 7853.981633 \\
+			0.000100 & 1000000 & 0.004453 & 7853.981617 \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\begin{table}[h]
+	\caption{mpirun --hostfile hosts.txt -np 2}
+	\begin{center}
+		\begin{tabular}{|l|l|l|l|}
+			\hline
+			Interval & Partitions & Time taken & Result \\
+			\hline
+			0.000001 & 100000000 & 0.240130 & 7853.981634 \\
+			0.000010 & 10000000 & 0.022352 & 7853.981633 \\
+			0.000100 & 1000000 & 0.002229 & 7853.981617 \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\begin{table}[h]
+	\caption{mpirun --hostfile hosts.txt -np 4}
+	\begin{center}
+		\begin{tabular}{|l|l|l|l|}
+			\hline
+			Interval & Partitions & Time taken & Result \\
+			\hline
+			0.000001 & 100000000 & 0.136689 & 7853.981634 \\
+			0.000010 & 10000000 & 0.011185 & 7853.981633 \\
+			0.000100 & 1000000 & 0.001120 & 7853.981617 \\
+			\hline
+		\end{tabular}
+	\end{center}
+\end{table}
+
+\begin{table}[h]
+	\caption{lmpirun --hostfile hosts.txt -np 8}
+	\begin{center}
+		\begin{tabular}{|l|l|l|l|}
+			\hline
+			Interval & Partitions & Time taken & Result \\
+			\hline
+			0.000001 & 100000000 & 0.116957 & 7853.981634 \\
+			0.000010 & 10000000 & 0.009895 & 7853.981633 \\
+			0.000100 & 1000000 & 0.000946 & 7853.981617 \\
+			\hline
+	\end{tabular}
+	\end{center}
+\end{table}
+
 	\section{Conclusion}
-	\par This is just a scratch on the surface of OpenMp please pay a visit to their website to learn more: \href{https://www.openmp.org/}{https://www.openmp.org/}.
+		\par The demonstrated example \textbf{does not imply} the superiority of OpenMP over Open MPI because, depending on the problem to be solved and the environment involved, one or the other may be more suitable for the task.
+
+		\par OpenMP is used when there is a need to utilize most of the local computer's resources for processing.
+OpenMP programs are simpler to code as well, thanks to the use of \textbf{\#pragma} compiler directives.
+
+		\par On the other hand, Open MPI is better suited when the problem needs to be divided into larger pieces and processed across multiple machines.
+However, it requires the programmer to split the problem, resulting in a more complex and error-prone code.
+
 		\par \textit{That is all folks!}

### 2023-07-11 code2/main.cpp

@@ -1,9 +1,7 @@
 #include <cmath>
-#include <cstddef>
 #include <iomanip>
 #include <iostream>
 #include <mpi.h>
-#include <ostream>
 
 inline long double f(long double x)
 {
@@ -68,7 +66,7 @@ inline void experiment(double lowerLimit, double upperLimit, double interval)
 
 int main(int argc, char *argv[])
 {
-    MPI_Init(NULL, NULL);
+    MPI_Init(&argc, &argv);
 
     double lowerLimit = 0, upperLimit = 100;
 

### 2023-07-11 code2/runmpi.sh

@@ -1,2 +1,2 @@
 # Change "build-IntegralEvaluator-Desktop-Release/parallel_integration" to the path of your open MPI executable
-mpirun --hostfile hosts.txt -np 10 ./build-IntegralEvaluator-Desktop-Release/parallel_integration
+mpirun --hostfile hosts.txt -np 1 ./build-IntegralEvaluator-Desktop-Release/parallel_integration

### 2023-07-12 documentation/article03/article.tex
Minor fixes

@@
 
-		\subsection{Open MPI}
+		\subsection{Open MPI discussion}
 		\par On line 4 there is a \textbf{required} include for the Open MPI library.
@@
 	\section{Conclusion}
-		\par The demonstrated example \textbf{does not imply} the superiority of OpenMP over Open MPI because, depending on the problem to be solved and the environment involved, one or the other may be more suitable for the task.
+		\par The demonstrated example \textbf{does not imply} the superiority of OpenMP over Open MPI or the other way around because, depending on the problem to be solved and the environment involved, one or the other may be more suitable for the task.
 
@@
 
+		\par Hope it helps and, for more information, please consult the references.\newline
+
 		\par \textit{That is all folks!}
